---
title: "Spring Data Compilation, Cleaning and Descriptive Statistics"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r include = F, read_in_data}
library(knitr)
library(tidyverse)
library(plotly)
knitr::opts_chunk$set(echo = F, message = F, warning = F)

```

```{r functions}
# Make function to append filtering step description, row count and count of unique pat ids.
add_filter_step = function(rct, dat, desc){
  append(
    rct,
    list(
      list(
      step = desc,
      row_count = nrow(dat),
      unique_pats = length(unique(dat$pat_id))
    )
    )
  )
}
```

```{r load_in_data_and_make_full_table}
# Import dataset.
treat = readr::read_csv("Data/Branch Care/Data/treatment_table_20260203 Adam Walter.csv") |> 
  dplyr::select(-pregnant)
str_egg = readr::read_csv("Data/Branch Care/Data/structured_egg_table_20260130 Adam Walter.csv")
raw_embryo_tbl = readr::read_csv("Data/Branch Care/Data/MLData_Nov25 Adam Walter/Embryo.csv")

# Remove extraneous columns in common with treatment table.
str_egg = str_egg |>
  dplyr::select(-c(attempt, age))

# Narrow columns of raw_embryo_tbl
raw_embryo_tbl = raw_embryo_tbl |> 
  dplyr::select(pat_id, egg_id, pregnant, donated, egg_maturity, pronuclei, thawed_oocyte, blast_rating, live_birth_Derive)

# Join structured egg tbl on with pat_id and tx_id
treat_w_egg = treat |> 
  dplyr::left_join(str_egg, by = dplyr::join_by(pat_id, tx_id))

treat_w_egg_and_embryo = treat_w_egg |> 
  dplyr::left_join(raw_embryo_tbl, by = dplyr::join_by(pat_id, egg_id))

# Initialize a list to track the rows we drop at each filtering step.
row_count_tbl = list()

# Append step 1's row count and description.
row_count_tbl = add_filter_step(row_count_tbl, treat_w_egg_and_embryo, "Cleaned treatment table from Branch with raw egg and embryo tables joined")

```

```{r}
# Prior to any filtering, these are the levels of the treatment variable.

# treat_w_egg_and_embryo |> 
#   dplyr::count(treatment, sort = T) |> 
#   knitr::kable()
```

```{r apply_filtering_steps}
# Remove "test" rows
treat_w_egg_and_embryo = treat_w_egg_and_embryo |> 
  dplyr::filter(pat_id != "C100")

row_count_tbl = add_filter_step(row_count_tbl, treat_w_egg_and_embryo, "Removed 'test' rows (e.g. pat_id of C100) (one row per egg)")

# # Check that we now have 'pregnant' column.
# if(interactive()){
#   treat_w_egg_and_embryo |> 
#   dplyr::count(pregnant)
# }


treat_w_egg_and_embryo = treat_w_egg_and_embryo |>
  dplyr::filter(treatment %in% c("IVF","ICSI"))

row_count_tbl = add_filter_step(row_count_tbl, treat_w_egg_and_embryo, "Filtered for just IVF or ICSI in 'treatment' column (one row per egg)")

# treat_w_egg_and_embryo = treat_w_egg_and_embryo |> 
#   dplyr::filter(pregnant %in% c('Clinical')) 
# 
# row_count_tbl = add_filter_step(row_count_tbl, treat_w_egg_and_embryo, "Filtered for just 'Clinical' in 'pregnant' column (one row per egg)")

grouping_vars = c('pat_id','tx_id','tx_cycle')

  # Add a categorical binned version of age.
treat_w_egg_and_embryo = treat_w_egg_and_embryo |>
  dplyr::mutate(age_bin = dplyr::case_when(
    age > 42 ~ ">42",
    age > 40 ~ "41-42",
    age > 37 ~ "38-40",
    age > 34 ~ "35-37",
    age < 35 ~ "<35",
    T ~ "Unknown"
  )) |> 
  dplyr::mutate(age_bin = factor(age_bin, levels = c("Unknown","<35","35-37","38-40","41-42",">42")))
```

```{r}
# treat_w_egg_and_embryo |> 
#   dplyr::filter(pat_id == 'C81763') |> 
#   dplyr::count(tx_id, attempt) |> 
#   knitr::kable()
```

```{r summarize_dataset_to_cld}
# Summarize to cycle-level table
cld <- treat_w_egg_and_embryo %>%
  # Filter for IVF and ICSI
  # dplyr::filter(treatment %in% c("IVF","ICSI")) |> 
  mutate(
    amh = coalesce(amh_start, amh),
    afc = coalesce(afc_first_baseline, afc_other)
  ) %>%
  group_by(pat_id, tx_id) %>%
  summarise(
    across(-c(pregnant, live_birth_Derive), first),
    pregnant_bool = as.integer(any(grepl("Clinical", pregnant, ignore.case = TRUE), na.rm = TRUE)),
    birth_bool = as.integer(any(grepl("Outcome Live Birth", live_birth_Derive, ignore.case = TRUE), na.rm = TRUE)),
    .groups = "drop"
  )

row_count_tbl = add_filter_step(row_count_tbl, cld, "Summarizing cycle-level table (one row per cycle)")

cld <- cld %>%
  select(dplyr::all_of(grouping_vars),
         attempt,
         treatment_start_date,
         treatment,
         primary_infertility_diagnosis,
         secondary_infertility_diagnosis,
         # number_eggs_collected,
         # number_m2_eggs,
         # number_2pn_eggs,
         # number_euploid_embryos,
         donated,
         num_eggs_collected,
         total_fertilized,
         num_m2_eggs,
         num_2pn_from_fresh_eggs,
         num_2pn_from_frozen_eggs,
         num_total_blast,
         num_pgds_done,
         num_euploid_blast,
         age,
         age_bin,
         bmi,
         amh,
         afc,
         pregnant_bool,
         birth_bool)

cld_2022 = cld |> 
  dplyr::filter(lubridate::year(treatment_start_date) == 2022) |> 
  dplyr::filter(!duplicated(tx_cycle))

# cld |> 
#   dplyr::count(pat_id, attempt, sort = T)

row_count_tbl = add_filter_step(row_count_tbl, cld_2022, "Filtering to treatments in 2022 (one row per cycle)")

```

# Data Filtering Steps {.tabset}

```{r show_filtering_row_counts}
row_count_tbl |> 
  dplyr::bind_rows() |> 
  knitr::kable()
```

## Cleaned Dataset (all years) {.tabset}

We took the cleaned treatment table provided by Adam Walter and left joined on the structured egg table and the embryo table. We then summarised key values from across this composite dataset to produce the following table:

```{r}
cld |> 
  DT::datatable()
```

## Diagnosis Levels (2022) {.tabset}

### Primary, Raw

```{r clean_up_infertility_diagnosis_column}
# cat("\nBefore cleaning up the primary_infertility_diagnosis column, it has these levels: ")

cld_2022 |> 
  dplyr::count(primary_infertility_diagnosis, sort = T) |> 
  DT::datatable()

cld_2022 = cld_2022 |> 
  dplyr::mutate(primary_infertility_diagnosis_cl = dplyr::case_when(
    stringr::str_detect(primary_infertility_diagnosis,'Tubal') ~ "Tubal factor",
    stringr::str_detect(primary_infertility_diagnosis,'Uterine') ~ "Uterine factor",
    stringr::str_detect(primary_infertility_diagnosis,'Male') ~ "Male factor",
    stringr::str_detect(primary_infertility_diagnosis,'Ovul') ~ "Ovulatory dysfunction",
    stringr::str_detect(primary_infertility_diagnosis,'[rR]eserve') ~ "Diminished ovarian reserve",
    stringr::str_detect(primary_infertility_diagnosis,'[bB]anking') ~ "Embryo Banking",
    is.na(primary_infertility_diagnosis) ~ "None",
    T ~ primary_infertility_diagnosis
  )) 
```

### Primary, Cleaned

```{r}
# cat("\nAfter cleaning, it has these levels: ")

cld_2022 |> 
  dplyr::count(primary_infertility_diagnosis_cl, sort = T) |> 
  DT::datatable()
```

### Secondary, Raw

```{r}

# cat("\nBefore cleaning up the secondary_infertility_diagnosis column, it has these levels: ")

cld_2022 |> 
  dplyr::count(secondary_infertility_diagnosis, sort = T) |> 
  DT::datatable()

cld_2022 = cld_2022 |> 
  dplyr::mutate(secondary_infertility_diagnosis_cl = dplyr::case_when(
    stringr::str_detect(secondary_infertility_diagnosis,'Tubal') ~ "Tubal factor",
    stringr::str_detect(secondary_infertility_diagnosis,'Uterine') ~ "Uterine factor",
    stringr::str_detect(secondary_infertility_diagnosis,'Male') ~ "Male factor",
    stringr::str_detect(secondary_infertility_diagnosis,'Ovul') ~ "Ovulatory dysfunction",
    stringr::str_detect(secondary_infertility_diagnosis,'[rR]eserve') ~ "Diminished ovarian reserve",
    stringr::str_detect(secondary_infertility_diagnosis,'[bB]anking') ~ "Embryo Banking",
    is.na(secondary_infertility_diagnosis) ~ "None",
    T ~ secondary_infertility_diagnosis
  )) 
```

### Secondary, Cleaned

```{r}

# cat("\nAfter cleaning, it has these levels: ")

cld_2022 |> 
  dplyr::count(secondary_infertility_diagnosis_cl, sort = T) |> 
  DT::datatable()

just_secondary = cld_2022 |> 
  dplyr::filter(is.na(primary_infertility_diagnosis_cl), !is.na(secondary_infertility_diagnosis_cl))

cat(paste0("\n",nrow(just_secondary)," patients have info for the secondary diagnosis but not the first."))

```

## ART table summaries {.tabset}

To compare with published ART tables for Spring Fertility here: https://art.cdc.gov/

### Cycle-level Summaries
```{r}

cld_2022 = cld_2022 |> 
  dplyr::mutate(pregnant_bool = ifelse(pregnant_bool == 1, "yes", "no"),
                birth_bool = ifelse(birth_bool == 1, "yes", "no"))

cld_2022 |> 
  dplyr::count(treatment) |> 
  knitr::kable()

cld_2022 |> 
  dplyr::count(treatment, pregnant_bool) |> 
  knitr::kable()

cld_2022 |> 
  dplyr::mutate(diagnoses_involve_adv_mat_age = primary_infertility_diagnosis_cl == "Advanced Maternal Age" | secondary_infertility_diagnosis_cl == "Advanced Maternal Age") |> 
  dplyr::count(diagnoses_involve_adv_mat_age) |> 
  knitr::kable()

cld_2022_sum_tbl = cld_2022 |> 
  dplyr::mutate(fresh_embryo = num_2pn_from_fresh_eggs > 0,
                frozen_embryo = num_2pn_from_frozen_eggs > 0) |> 
  dplyr::count(donated,fresh_embryo,frozen_embryo,pregnant_bool,birth_bool)

cld_2022_sum_tbl |> 
  knitr::kable()

cld_2022_sum_tbl |> 
  dplyr::count(fresh_embryo, frozen_embryo, pregnant_bool, wt = n) |> 
  knitr::kable()

cld_2022 |> 
  dplyr::count(birth_bool) |> 
  knitr::kable()
# cld_2022 %>%
#   count(secondary_infertility_diagnosis)
# 

# 
# cld %>%
#   count(year(treatment_start_date)) |> knitr::kable()

# cld_2022 %>%
# distinct(pat_id) |> 
#   knitr::kable()
```

### Age of Patient Histogram
```{r}
p = cld_2022 |> 
  dplyr::group_by(pat_id) |> 
  dplyr::slice(1) |> 
  dplyr::ungroup() |> 
  dplyr::mutate(total_patients = dplyr::n()) |> 
  dplyr::group_by(age_bin) |> 
  dplyr::reframe(patients_in_bin = dplyr::n(), total_patients = unique(total_patients)) |> 
  dplyr::ungroup() |> 
  dplyr::mutate(perc_in_bin = round(patients_in_bin / total_patients,2)) |> 
  ggplot() + 
  geom_col(aes(x = age_bin, y = perc_in_bin)) + 
  labs(y = "Percent (%)", x = "Age of Patient") +
  scale_y_continuous(labels = scales::percent)

ggplotly(p)
```

```{r}

# #### Success Rates: Patients Using Own Eggs
# 
# # NOTE: I am unsure if I've used the correct variable to infer if patients are using their own eggs...
# 
# srdat = cld_2022 |> 
#   dplyr::group_by(age_bin, primary_infertility_diagnosis_cl) |> 
#   dplyr::reframe(treatments_with_1_plus_eggs_collected = sum(num_eggs_collected >= 1),
#                  treatments_leading_to_live_birth = sum(birth_bool == 'yes',na.rm=T)) |> 
#   dplyr::mutate(treatment_success_perc = treatments_leading_to_live_birth / treatments_with_1_plus_eggs_collected) |>
#   dplyr::mutate(treatment_success_perc = tidyr::replace_na(treatment_success_perc, 0))
#   
# srdat |> 
#   ggplot() + 
#   geom_col(aes(x = age_bin, y = treatment_success_perc), linewidth = 2) +
#   facet_wrap( ~ primary_infertility_diagnosis_cl, scales = 'free') + 
#   scale_y_continuous(labels = scales::percent) + 
#   labs(y = 'Success Rate (%)', x = "Age Category")
```

```{r}

```

