---
title: "Insurance Modeling Analysis"
author: "Charm Economics"
date: "`r Sys.Date()`"
output: 
  rmdformats::downcute
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(dplyr)
library(ggplot2)
library(VGAM)
library(pROC)
library(MASS)
library(ResourceSelection)
library(knitr)
library(tidyr)
library(readr)
set.seed(123)
```

```{r define_functions}
source(file.path(here::here(),"R/utils/modeling_stats_utility_functions.R"))
```

```{r read_in_and_clean_cycle_data, include=F}
df <- readr::read_rds("data/clean_data/cld_filtered.rds") |> 
  dplyr::rename(pregnant = pregnant_bool)

egg_df = readr::read_rds("data/clean_data/cleaned_egg_level_df.rds")

# Number of unique PAT ids
df = df |> dplyr::mutate(the_year = lubridate::year(treatment_start_date))
```

# Descriptive Statistics {.tabset}

## Number of Unique Patients

```{r}
df |> 
  dplyr::select(pat_id, age) |> 
  dplyr::distinct() |> 
  dplyr::count(age) |> 
  knitr::kable()
```

## Age, AMH and AFC Distributions
```{r}
suppressWarnings(
df |> 
  dplyr::select(age, afc, amh) |> 
  tidyr::pivot_longer(cols = dplyr::everything()) |> 
  dplyr::mutate(name = factor(name, levels = c('age','afc','amh'))) |> 
  ggplot(aes(value)) +
  facet_wrap(~ name, scales = 'free', ncol = 1) +
  geom_histogram(bins = 20) + 
  scale_x_continuous(breaks = scales::breaks_pretty())
)

df |> 
  dplyr::select(age, afc, amh) |> 
  tidyr::pivot_longer(cols = dplyr::everything()) |> 
  dplyr::group_by(name) |> 
  dplyr::reframe(maximum = max(value,na.rm=T), minimum = min(value, na.rm=T), median = median(value, na.rm=T)) |> 
  knitr::kable()
```


## Descriptive Pregnancy Rates
```{r}
pregnancy_rates <- df %>%
  group_by(age) %>%
  summarize(
    cycles = n(),
    pregnancies = sum(pregnant),
    pregnancy_rate = pregnancies / cycles
  )

kable(pregnancy_rates, digits = 3)

ggplot(pregnancy_rates, aes(x = age, y = pregnancy_rate)) +
  geom_col() +
  labs(
    title = "Pregnancy Rate by Age Group",
    x = "Age Group",
    y = "Pregnancy Rate"
  )
```

## Dataset Date Range
```{r}
df |> 
  dplyr::reframe(earliest_date = min(treatment_start_date, na.rm=T),
                 latest_date = max(treatment_start_date, na.rm=T)) |> 
  knitr::kable()

df |> 
  dplyr::mutate(year = lubridate::year(treatment_start_date)) |> 
  dplyr::count(year, name = 'unique patient/treatment cycle combinations')

ggplot(df) + 
  geom_histogram(aes(treatment_start_date)) + 
  scale_x_date(date_breaks = 'year', date_labels = '%Y') +
  theme_minimal()
```

## Key Variables
```{r make_tbl_for_key_variables}
var_summary <- df %>%
  mutate(
    total_2pn_eggs = num_2pn_from_fresh_eggs + num_2pn_from_frozen_eggs
  ) %>%
  group_by(age) %>%
  summarise(
    n_cycles = n(),
    
    # Euploid embryos
    mean_euploid    = mean(num_euploid_blast, na.rm = TRUE),
    iqr_euploid_l   = quantile(num_euploid_blast, 0.25, na.rm = TRUE),
    iqr_euploid_u   = quantile(num_euploid_blast, 0.75, na.rm = TRUE),
    pct_ge1_euploid = mean(coalesce(num_euploid_blast, 0) >= 1) * 100,
    
    # Eggs retrieved
    mean_eggs       = mean(num_eggs_collected, na.rm = TRUE),
    iqr_eggs_l      = quantile(num_eggs_collected, 0.25, na.rm = TRUE),
    iqr_eggs_u      = quantile(num_eggs_collected, 0.75, na.rm = TRUE),
    
    # MII eggs
    mean_m2         = mean(num_m2_eggs, na.rm = TRUE),
    iqr_m2_l        = quantile(num_m2_eggs, 0.25, na.rm = TRUE),
    iqr_m2_u        = quantile(num_m2_eggs, 0.75, na.rm = TRUE),
    
    # Fertilized (2PN)
    mean_2pn        = mean(total_2pn_eggs, na.rm = TRUE),
    iqr_2pn_l       = quantile(total_2pn_eggs, 0.25, na.rm = TRUE),
    iqr_2pn_u       = quantile(total_2pn_eggs, 0.75, na.rm = TRUE),
    
    # AMH
    mean_amh        = mean(amh, na.rm = TRUE),
    iqr_amh_l       = quantile(amh, 0.25, na.rm = TRUE),
    iqr_amh_u       = quantile(amh, 0.75, na.rm = TRUE),
    
    # AFC
    mean_afc        = mean(afc, na.rm = TRUE),
    iqr_afc_l       = quantile(afc, 0.25, na.rm = TRUE),
    iqr_afc_u       = quantile(afc, 0.75, na.rm = TRUE),
    
    .groups = "drop"
  )


var_summary_fmt <- var_summary %>%
  mutate(
    eggs_fmt = sprintf("%.1f (%.1f–%.1f)", mean_eggs, iqr_eggs_l, iqr_eggs_u),
    m2_fmt   = sprintf("%.1f (%.1f–%.1f)", mean_m2, iqr_m2_l, iqr_m2_u),
    p2n_fmt  = sprintf("%.1f (%.1f–%.1f)", mean_2pn, iqr_2pn_l, iqr_2pn_u),
    amh_fmt  = sprintf("%.2f (%.2f–%.2f)", mean_amh, iqr_amh_l, iqr_amh_u),
    afc_fmt  = sprintf("%.1f (%.1f–%.1f)", mean_afc, iqr_afc_l, iqr_afc_u),
    euploid_fmt = sprintf("%.1f (%.1f–%.1f)", mean_euploid, iqr_euploid_l, iqr_euploid_u),
    euploid_pct = sprintf("%.1f%%", pct_ge1_euploid)
  )

var_summary_fmt |> 
  knitr::kable()
```

# Removal of Patients over 43 years of age

```{r}
df = df |> 
  dplyr::filter(age <= 43)

egg_df = egg_df |> 
  dplyr::filter(age <= 43)
```

# Patient-level Statistical Models {.tabset}

Two models here:

Live birth model
Euploid embryo model


```{r make_model_1_dataset}

# <!-- This filter reduces the number of rows of data in this cycle-level table from `r nrow(df)` to `r nrow(m1df)`. -->
  
# Note - models in this section include only treatments of type 'ICSI' (the majority of the treatments) and NOT those of type 'IVF'.

euploid_embryos_from_egg_df = egg_df |> 
  dplyr::select(pat_id,tx_id,tx_cycle,pgd_clinresult) |> 
  dplyr::group_by(pat_id,tx_id) |> 
  dplyr::reframe(euploid_success = sum(pgd_clinresult %in% c("Euploid", "Normal") & !is.na(pgd_clinresult)) > 0)

m1df = df |>
  dplyr::left_join(euploid_embryos_from_egg_df) |> 
  dplyr::arrange(pat_id, treatment_start_date) |> 
  dplyr::group_by(pat_id) |> 
  dplyr::mutate(cycle_number = dplyr::row_number()) |> 
  dplyr::ungroup() |> 
  dplyr::select(pat_id, euploid_success, age, afc, amh, birth_bool, cycle_number, treatment_start_date) |> 
  tidyr::drop_na()

  # dplyr::filter(treatment == 'ICSI')
```



## Live-birth Model {.tabset}

### Model Outputs

```{r run_first_two_models}

# <!-- After dropping rows with NA for any of `euploid_success`, `age`, `afc`, or `amh`, the treatment cycle-level dataset has `r nrow(m1df)` observations, `r nrow(m1df_eup)` of which achieved a euploid embryo success. -->
  
euploid_model <- glm(
  euploid_success ~ age + amh + afc,
  family = binomial,
  data = m1df
)

live_birth_model <- glm(
  birth_bool ~ age + amh + afc,
  family = quasibinomial,
  data = m1df |> dplyr::filter(euploid_success)
)

m1df$pred_euploid <- predict(euploid_model, new_data = m1df, type = "response")

m1df_eup = m1df |> 
  dplyr::filter(euploid_success)

m1df_eup$pred_lb_given_euploid <- predict(
  live_birth_model,
  newdata = m1df_eup,
  type = "response"
)

m1df_eup$pred_live_birth <- m1df_eup$pred_euploid * m1df_eup$pred_lb_given_euploid

livebirth_sim_results <- m1df_eup |>
  dplyr::rowwise() |>
  dplyr::mutate(
    success_sim = rbinom(1, size = 1, prob = pred_live_birth),
    livebirth_success_simulated = success_sim >= 1
  ) |>
  dplyr::ungroup()
```

<!-- The live-birth model's measure of dispersion is `r check_dispersion(live_birth_model)`. -->

```{r}
livebirth_sim_results |>
  DT::datatable()

livebirth_sim_results |> 
  dplyr::count(livebirth_success_simulated) |> 
  dplyr::mutate(percentage = round( 100 * n / sum(n), 2)) |> 
  knitr::kable()
```

```{r read_in_egg_dataframe_and_clean_it, include = F}

# I'm unsure if we NEED to read in a new copy of the structured egg table... but if we do, we follow the same cleaning steps as above.


# Zoom in on relevant rows, drop rows with no info for age and attempt.
egg_df = egg_df |> 
  dplyr::select(pat_id, egg_id, age, attempt, pgd_clinresult, frozen_embryo, frozen_egg, afc_other, afc_first_baseline, amh, amh_start) |> 
  dplyr::filter(!is.na(age))

# Add treatment course ID and cycle number ID columns
df = df |> 
  dplyr::arrange(pat_id, treatment_start_date) |>
  dplyr::group_by(pat_id) |>
  dplyr::mutate(
    gap_days = as.numeric(treatment_start_date - dplyr::lag(treatment_start_date)),
    new_treat_course = ifelse(is.na(gap_days) | gap_days > 365, 1, 0),
    treat_course_id = cumsum(new_treat_course)
  ) |> 
  dplyr::select(-c(gap_days,new_treat_course)) |> 
  # Define cycle number within a treatment course
  dplyr::group_by(pat_id, treat_course_id) |>
  dplyr::arrange(treatment_start_date) |>
  # Cycle number is the key column we're making here - within a given 'treat_course', 
  # we can have various related cycles (cycle 1, 2, etc.)
  dplyr::mutate(cycle_number = dplyr::row_number()) |> 
  dplyr::ungroup()

cost_per_cycle <- 22014
payout_amount <- 17611
risk_margin <- 0.20
admin_load <- 0.10

n_sim <- 5000 # Monte Carlo iterations

```

### Multi-cycle (Cumulative) Probabilities
```{r}
# --- Closed-form cumulative probability (Live Birth) ---
livebirth_multi_cycle = m1df_eup |> 
 dplyr::group_by(pat_id) |>
  dplyr::summarise(
    p1 = first(pred_live_birth)
  ) |>
  dplyr::mutate(
    cycle_1 = 1 - (1 - p1)^1,
    cycle_2 = 1 - (1 - p1)^2,
    cycle_3 = 1 - (1 - p1)^3,
    failure_1 = (1 - p1)^1,
    failure_2 = (1 - p1)^2,
    failure_3 = (1 - p1)^3
  )

livebirth_multi_cycle |> 
  DT::datatable()

readr::write_csv(livebirth_multi_cycle, "outputs/livebirth_multicycle_probabilities.csv")

```

### Simulated Probabilities averaged by Age
```{r}
livebirth_sim_results |> 
  dplyr::group_by(age) |> 
  dplyr::reframe(
    pct_livebirth_simulated = mean(livebirth_success_simulated, na.rm = TRUE)
  ) |> 
  dplyr::arrange(age) |> 
  knitr::kable()
```

## Euploid Embryo Model (Binomial) {.tabset}

### Model Outputs

<!-- The euploid model's measure of dispersion is `r check_dispersion(euploid_model)`. -->

```{r}
euploid_sim_results <- m1df |>
  dplyr::rowwise() |>
  dplyr::mutate(
    success_sim = rbinom(1, size = 1, prob = pred_euploid),
    ge1_euploid_sim = success_sim >= 1
  ) |>
  dplyr::ungroup()

euploid_sim_results |>
  DT::datatable()

euploid_sim_results |> 
  dplyr::count(ge1_euploid_sim) |> 
  dplyr::mutate(percentage = round( 100 * n / sum(n), 2)) |> 
  knitr::kable()
```

### Multi-cycle (Cumulative) Probabilities
```{r}
# --- Closed-form cumulative probability (Live Birth) ---
euploid_multi_cycle = m1df |> 
 dplyr::group_by(pat_id) |>
  dplyr::summarise(
    p1 = first(pred_euploid)
  ) |>
  dplyr::mutate(
    cycle_1 = 1 - (1 - p1)^1,
    cycle_2 = 1 - (1 - p1)^2,
    cycle_3 = 1 - (1 - p1)^3,
    failure_1 = (1 - p1)^1,
    failure_2 = (1 - p1)^2,
    failure_3 = (1 - p1)^3
  )

euploid_multi_cycle |> 
  DT::datatable()

readr::write_csv(euploid_multi_cycle, "outputs/euploid_embryo_multicycle_probabilities.csv")

```

### Simulated Probabilities averaged by Age
```{r}
euploid_sim_results |> 
  dplyr::group_by(age) |> 
  dplyr::reframe(
    pct_1_plus_euploid_sim = mean(ge1_euploid_sim, na.rm = TRUE)
  ) |> 
  dplyr::arrange(age) |> 
  knitr::kable()
```

# Egg Actuarial Models (Beta-binomial / Binomial + Monte-Carlo Simulation) {.tabset}

## Single-cycle Egg Model (Binomial model) {.tabset}

```{r}
# Remove rows where amh is unknown.
df = df |> dplyr::filter(!is.na(amh))
```

### Model Outputs

Note: this model uses the patient / cycle-level table and does not currently filter away rows for eggs of cryo_day_2 or cryo_day_3; N of this dataframe is therefore `r nrow(df)`.

```{r}

df_no_na = df |> dplyr::select(pat_id, tx_id, num_eggs_collected, age, amh, afc) |> tidyr::drop_na()

# First, construct a Poisson-distribution of the number of eggs collected based on the data we have
egg_model <- MASS::glm.nb(
  num_eggs_collected ~ age + amh + afc,
  # family = quasipoisson(link = "log"),
  data = df_no_na
)

df_no_na$pred_eggmodel_1 = predict(egg_model, new_data = df_no_na, type = "response")

# Next, for each patient, based on their characteristics (age, afc, amh), simulate 
# 2000 hypothetical egg gathering treatments; these make up the simulated
# expectable distribution that is unique to that patient. We then take 
# out the 90th and 95th percentile of expectable eggs and use that as a 
# threshold for whether or not a simulated session would be successful
# for this patient. It's all about likelihoods!
egg_1cycle_simulated_thresholds_by_patient = df_no_na |> 
  dplyr::group_by(pat_id) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    sim_egg_1cycle_distribution = simulate_eggs(.x)
    sim_egg_1cycle_90 = quantile(sim_egg_1cycle_distribution, 0.1)
    sim_egg_1cycle_95 = quantile(sim_egg_1cycle_distribution, 0.05)
    output = data.frame(
      pat_id = .x$pat_id,
      sim_egg_1cycle_90,
      sim_egg_1cycle_95
    )
    row.names(output) = NULL
    output |> tidyr::as_tibble()
  },.progress) |> 
  dplyr::bind_rows() |> 
  dplyr::distinct() |> 
  suppressWarnings()

df_with_thresholds <- df_no_na |>
  dplyr::left_join(egg_1cycle_simulated_thresholds_by_patient, by = "pat_id")

# Now we simulate a single 'draw' for each patient, compare that draw to the 
# patient-specific distribution of expectable number of eggs, and classify
# the draw as either a success or a failure (i.e. above or below the threshold, respectively)
eggmodel_sim_results <- df_with_thresholds |>
  dplyr::rowwise() %>%
  dplyr::mutate(
    sim_eggs_one_cycle = simulate_eggs(., n = 1),
    success_90 = sim_eggs_one_cycle >= sim_egg_1cycle_90,
    success_95 = sim_eggs_one_cycle >= sim_egg_1cycle_95
  ) |>
  dplyr::ungroup()

eggmodel_sim_results |> 
  dplyr::count(success_90, success_95) |> 
  dplyr::mutate(percentage = round(100*n/sum(n),4))

# Conversion of simulated successes/failures to predicted probabilities so we can
# apply the same model tests and validations to the egg models as we do to 
# the live-birth and euploid embryo eggs.
eggmodel_sim_results <- eggmodel_sim_results %>%
  dplyr::mutate(
    pred_prob_success_90 = estimate_prob_success(
      ., simulate_eggs, "sim_egg_1cycle_90", n_sim = 1000
    ),
    pred_prob_success_95 = estimate_prob_success(
      ., simulate_eggs, "sim_egg_1cycle_95", n_sim = 1000
    )
  )
```

### Simulated Probabilities averaged by Age
```{r}
eggmodel_sim_results |>
  dplyr::group_by(age) |>
  dplyr::reframe(
    pct_success_90 = mean(success_90, na.rm = TRUE),
    pct_success_95 = mean(success_95, na.rm = TRUE),
  ) |>
  dplyr::arrange(age) |>
  knitr::kable()
```

## Two-cycle Egg Model (Binomial model) {.tabset}

Note: this model includes only patients that have more than one treatment recorded in the dataset.
Note: this model assumes that treatments that occur within 365 days belong to a single treatment course, i.e. a collection of cycles including cycle 1, cycle 2, etc.; this assumption was necessary to compute this model because the dataset lacks a variable that otherwise links treatments across time. We could modify this assumption by removing this 365-day date proximity threshold, though this may incorrectly imply relatedness between distinct treatments performed by a patient across multiple years.

### Model Outputs

```{r}
df_2 <- df %>%
  dplyr::arrange(pat_id, treatment_start_date) %>%
  dplyr::group_by(pat_id) %>%
  dplyr::mutate(
    eggs_cycle1 = num_eggs_collected,
    eggs_cycle2 = dplyr::lead(num_eggs_collected),
    eggs_2cycle_total = eggs_cycle1 + eggs_cycle2
  ) %>%
  dplyr::filter(!is.na(eggs_2cycle_total)) %>%
  dplyr::ungroup()

# Ensure we don't have missing data for any of our variables.
df_2 = df_2 |> 
  dplyr::select(pat_id, eggs_2cycle_total, age, amh, afc, eggs_cycle1, eggs_cycle2, eggs_2cycle_total) |> 
  tidyr::drop_na()

egg2_model <- MASS::glm.nb(
  eggs_2cycle_total ~ age + amh + afc + eggs_cycle1,
  # family = quasipoisson(link = "log"),
  data = df_2
)

df_2$pred_eggmodel_2 = predict(egg2_model, new_data = df_2, type = "response")

# Next, for each patient, based on their characteristics (age, afc, amh), simulate 
# 2000 hypothetical egg gathering treatments; these make up the simulated
# expectable distribution that is unique to that patient. We then take 
# out the 90th and 95th percentile of expectable eggs and use that as a 
# threshold for whether or not a simulated session would be successful
# for this patient. It's all about likelihoods!
egg_2cycle_simulated_thresholds_by_patient = df_2 |> 
  dplyr::group_by(pat_id) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    sim_egg_2cycle_distribution = simulate_eggs2(.x)
    sim_egg_2cycle_90 = quantile(sim_egg_2cycle_distribution, 0.1)
    sim_egg_2cycle_95 = quantile(sim_egg_2cycle_distribution, 0.05)
    output = data.frame(
      pat_id = .x$pat_id,
      sim_egg_2cycle_90,
      sim_egg_2cycle_95
    )
    row.names(output) = NULL
    output |> tidyr::as_tibble()
  },.progress) |> 
  dplyr::bind_rows() |> 
  dplyr::distinct() |> 
  suppressWarnings()

df2_with_thresholds <- df_2 |>
  dplyr::left_join(egg_2cycle_simulated_thresholds_by_patient, by = "pat_id")

# Now we simulate a single 'draw' for each patient, compare that draw to the 
# patient-specific distribution of expectable number of eggs, and classify
# the draw as either a success or a failure (i.e. above or below the threshold, respectively)
eggmodel2_sim_results <- df2_with_thresholds |>
  dplyr::rowwise() %>%
  dplyr::mutate(
    sim_eggs_two_cycle = simulate_eggs2(., n = 1),
    success_90 = sim_eggs_two_cycle >= sim_egg_2cycle_90,
    success_95 = sim_eggs_two_cycle >= sim_egg_2cycle_95
  ) |>
  dplyr::ungroup()

eggmodel2_sim_results |> 
  dplyr::count(success_90, success_95) |> 
  dplyr::mutate(percentage = round(100*n/sum(n),4))

eggmodel2_sim_results <- eggmodel2_sim_results %>%
  dplyr::mutate(
    pred_prob_success_90 = estimate_prob_success(
      ., simulate_eggs2, "sim_egg_2cycle_90", n_sim = 1000
    ),
    pred_prob_success_95 = estimate_prob_success(
      ., simulate_eggs2, "sim_egg_2cycle_95", n_sim = 1000
    )
  )
```

### Simulated Probabilities averaged by Age
```{r}
eggmodel2_sim_results |>
  dplyr::group_by(age) |>
  dplyr::reframe(
    pct_success_90 = mean(success_90, na.rm = TRUE),
    pct_success_95 = mean(success_95, na.rm = TRUE),
  ) |>
  dplyr::arrange(age) |>
  knitr::kable()
```

```{r, include = F}
# Euploid model ROC / AUC
roc_livebirth <- roc(livebirth_sim_results$livebirth_success_simulated, livebirth_sim_results$pred_live_birth)
auc_livebirth <- auc(roc_livebirth)
roc_euploid <- roc(m1df$euploid_success, m1df$pred_euploid)
auc_euploid <- auc(roc_euploid)
roc_egg1 <- roc(eggmodel_sim_results$success_90, eggmodel_sim_results$pred_eggmodel_1)
auc_egg1 <- auc(roc_egg1)
roc_egg2 <- roc(eggmodel2_sim_results$success_90, eggmodel2_sim_results$pred_eggmodel_2)
auc_egg2 <- auc(roc_egg2)

# --- INTERPRETATION GUIDE ---
# AUC < 0.60 → Poor model (not usable)
# AUC 0.60–0.70 → Weak (needs improvement or more predictors)
# AUC 0.70–0.80 → Acceptable
# AUC 0.80–0.90 → Strong
# AUC > 0.90 → Possibly overfit (check carefully)

# plot(roc_euploid)

# Dispersion checks
disp_livebirth = check_dispersion(live_birth_model)
disp_euploid <- check_dispersion(euploid_model)
disp_egg <- check_dispersion(egg_model)
disp_egg2 <- check_dispersion(egg2_model)
```

# Base Model Validation
```{r}
validation_summary <- data.frame(
  Model = c("Live Birth","Euploid", "Egg 1 Cycle", "Egg 2 Cycle"),
  AUC = c(auc_livebirth, auc_euploid, auc_egg1, auc_egg2),
  Dispersion = c(disp_livebirth, disp_euploid, disp_egg, disp_egg2)
) |> 
  dplyr::mutate(interaction_term = 'None')

validation_summary |> knitr::kable()
```

# Models Including Interaction Terms
```{r}
# loop over potential models.
my_formulas = c(
  "base" = "age + amh + afc",
  "amh*afc" = "age + amh + afc + amh * afc",
  "age*afc" = "age + amh + afc + age * afc",
  "age*amh" = "age + amh + afc + amh * age",
  "age*amh+age*afc" = "age + amh + afc + amh * age + afc * age",
  "age*afc+amh*afc" = "age + amh + afc + afc * age + afc * amh",
  "age*amh+amh*afc" = "age + amh + afc + amh * age + afc * amh",
  "age*amh*afc" = "age + amh + afc + amh * age * afc"
  )

model_skeleton_list = list(
      
      # --- Euploid ---
      euploid = list(
        data = m1df,
        formula = euploid_success ~ .x,
        family = binomial,
        family_text = 'binomial',
        outcome = "euploid_success"
      ),
      
      # --- Live birth ---
      livebirth = list(
        data = m1df_eup,
        formula = birth_bool ~ .x,
        family = quasibinomial,
        family_text = 'quasibinomial',
        outcome = "birth_bool"
      ),
      
      egg1 = list(
        data = df_no_na,
        formula = num_eggs_collected ~ .x,
        family = quasipoisson(link = "log"),
        family_text = "egg",
        outcome = "num_eggs_collected"
      ),
      
      egg2 = list(
        data = df_2,
        formula = eggs_2cycle_total ~ .x,
        family = quasipoisson(link = "log"),
        family_text = "egg",
        outcome = "eggs_2cycle_total"
      )
    )

model_specs <- purrr::map(my_formulas, \(x){
  purrr::map(model_skeleton_list, \(y){
    dep_var = y$formula |> as.character()
    dep_var = dep_var[2]
    y$formula = paste0(dep_var," ~ ",x)
    y
  })
})

# Apply model specs list! Let's see what the results look like.

model_results = model_specs |> 
  purrr::list_flatten() |> 
  purrr::imap(function(spec, name) {
  
    spec = purrr::list_flatten(spec)
    
  if(spec$family_text == "egg"){
    model = MASS::glm.nb(spec$formula, data = spec$data) |> suppressWarnings()
  } else{
    model <- glm(spec$formula, data = spec$data, family = spec$family) |> suppressWarnings()
  }
  
  dispersion = check_dispersion(model)
  
  # Only compute AUC/Brier if binomial-type outcome
  is_binary <- spec$family_text %in% c("binomial", "quasibinomial")
  is_egg <- spec$family_text == "egg"
  
  metrics <- NULL
  
  if (is_egg) {
    
    data <- spec$data
    data$pred <- predict(model, newdata = data, type = "response")
    
    # Re-run simulation using THIS model
    sim_results <- data |>
      dplyr::rowwise() %>%
      dplyr::mutate(
        sim_val = simulate_eggs(., n = 1),   # or simulate_eggs2
        success = sim_val >= quantile(simulate_eggs(.), 0.1)
      ) |>
      dplyr::ungroup()
    
    # Estimate probabilities
    sim_results <- sim_results %>%
      dplyr::mutate(
        pred_prob = estimate_prob_success(
          ., simulate_eggs, "success", n_sim = 500
        )
      )
    
    metrics <- get_metrics_from_probs(
      observed = sim_results$success,
      predicted = sim_results$pred_prob
    )
    
  } else if (is_binary) {
    metrics <- get_metrics(model, spec$data, spec$outcome)
  } else {
    metrics <- c(AIC = AIC(model))
  }
  
  cv <- if (is_binary) {
    cv_evaluate(spec$formula, spec$data, spec$outcome, spec$family)
  } else {
    NULL
  }
  
  metrics['Dispersion'] = dispersion[[1]]
  
  list(
    model = model,
    metrics = metrics,
    cv_sd = if (!is.null(cv)) sd(cv$AUC, na.rm = TRUE) else NA
  )
},.progress=T) |> 
  suppressMessages()

model_summary <- purrr::imap_dfr(model_results, function(res, name) {
  tibble::tibble(
    # Flip first and last bits of model name
    model = paste0(stringr::str_extract(name,"(?<=_).*"), "_", stringr::str_extract(name,".*(?=_)")) |> 
      # Oop, Markdown is converting * to italics on occasion... swap those out and put ' x '
      stringr::str_replace_all("\\*", " x "),
    AUC = res$metrics["AUC"] %||% NA,
    Brier = res$metrics["Brier"] %||% NA,
    Cal_Slope = res$metrics["Cal_Slope"] %||% NA,
    CV_AUC_SD = res$cv_sd,
    AIC = res$metrics["AIC"] %||% NA,
    Dispersion = res$metrics["Dispersion"] %||% NA
  )
})

model_summary <- model_summary %>%
  dplyr::left_join(
    validation_summary %>% 
      dplyr::select(Model, Dispersion) |> 
      dplyr::mutate(model = snakecase::to_snake_case(Model)) |> 
      dplyr::select(-Model)
  ) |> 
  dplyr::arrange(model)

model_summary |> 
  knitr::kable()
```

# Test of Negative Binomial for Egg Models {.tabset}

```{r}

eg_patient = tidyr::tibble(
  age = 35,
  amh = mean(df_no_na$amh,na.rm=T),
  afc = mean(df_no_na$afc,na.rm=T)
) |> 
  mutate(eggs_cycle1 = round(age * -0.3 + amh * 1.5 + afc * 0.4 + 10))

n_sims = 5000

compare_thresholds <- function(mu, theta_new, theta_old, n = n_sims) {
  sims_new <- rnbinom(n, mu = mu, size = theta_new)
  sims_old <- rnbinom(n, mu = mu, size = theta_old)

  data.frame(
    t90_new  = quantile(sims_new, 0.10),
    t95_new  = quantile(sims_new, 0.05),
    t90_old  = quantile(sims_old, 0.10),
    t95_old  = quantile(sims_old, 0.05),
    t90_diff = quantile(sims_new, 0.10) - quantile(sims_old, 0.10),
    t95_diff = quantile(sims_new, 0.05) - quantile(sims_old, 0.05),
    mean_new = mean(sims_new),
    mean_old = mean(sims_old)
  )
}
```

## Egg Model 1

```{r egg_model_1_effect_of_altering_theta}
theta_egg1 = egg_model$theta
theta_old = 5

print(paste0("Model Theta: ",round(theta_egg1,4)))
print(paste0("Model Theta SE: ",round(theta_old,4)))

eg_patient %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    mu_egg1  = predict(egg_model,  newdata = cur_data(), type = "response"),
    egg1     = list(compare_thresholds(mu_egg1, theta_egg1, theta_old))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(age, amh, afc, mu_egg1, egg1) %>%
  tidyr::unnest_wider(egg1) %>%
  dplyr::mutate(model = "Egg 1-Cycle") |> 
  suppressWarnings() |> 
  knitr::kable()
```

## Egg Model 2

```{r egg_model_2_effect_of_altering_theta}
theta_egg2 = egg2_model$theta
theta_old = 5

print(paste0("Model Theta: ",round(theta_egg2,4)))
print(paste0("Model Theta SE: ",round(theta_old,4)))

eg_patient %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    mu_egg2  = predict(egg2_model,  newdata = cur_data(), type = "response"),
    egg2     = list(compare_thresholds(mu_egg2, theta_egg2, theta_old))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(age, amh, afc, mu_egg2, egg2) %>%
  tidyr::unnest_wider(egg2) %>%
  dplyr::mutate(model = "Egg 2-Cycle") |> 
  suppressWarnings() |> 
  knitr::kable()
```

# Comparison Plots for fixed versus fitted theta values {.tabset}
```{r comparison_plots_for_egg_models_theta_values}

mu1 <- predict(egg_model,  newdata = eg_patient, type = "response")
mu2 <- predict(egg2_model, newdata = eg_patient, type = "response")

plot_comparison <- function(mu, theta_new, theta_old, title_str, n = n_sims) {

  sims_new <- rnbinom(n, mu = mu, size = theta_new)
  sims_old <- rnbinom(n, mu = mu, size = theta_old)

  t90_new <- quantile(sims_new, 0.10)
  t95_new <- quantile(sims_new, 0.05)
  t90_old <- quantile(sims_old, 0.10)
  t95_old <- quantile(sims_old, 0.05)

  plot_df <- bind_rows(
    data.frame(eggs = sims_new, version = sprintf("Fitted theta = %.2f", theta_new)),
    data.frame(eggs = sims_old, version = sprintf("Original theta = %.1f", theta_old))
  )

  vlines <- data.frame(
    xint    = c(t90_new, t95_new, t90_old, t95_old),
    version = c(
      sprintf("Fitted theta = %.2f", theta_new),
      sprintf("Fitted theta = %.2f", theta_new),
      sprintf("Original theta = %.1f", theta_old),
      sprintf("Original theta = %.1f", theta_old)
    ),
    label = c(
      sprintf("90%% (%d eggs)", t90_new),
      sprintf("95%% (%d eggs)", t95_new),
      sprintf("90%% (%d eggs)", t90_old),
      sprintf("95%% (%d eggs)", t95_old)
    ),
    lty = c("90%", "95%", "90%", "95%")
  )

  ggplot(plot_df, aes(x = eggs, fill = version)) +
    geom_histogram(
      aes(y = after_stat(density)),
      binwidth = 1,
      alpha    = 0.5,
      position = "identity",
      color    = "white",
      linewidth = 0.2
    ) +
    geom_vline(
      data     = vlines,
      aes(xintercept = xint, color = version, linetype = lty),
      linewidth = 0.9
    ) +
    geom_text(
      data  = vlines,
      aes(x = xint, y = Inf, label = label, color = version),
      angle = 90, vjust = -0.4, hjust = 1.1, size = 3,
      inherit.aes = FALSE
    ) +
    scale_linetype_manual(
      name   = "Threshold",
      values = c("90%" = "dashed", "95%" = "dotted")
    ) +
    scale_fill_manual(values  = c("#378ADD", "#D85A30")) +
    scale_color_manual(values = c("#185FA5", "#993C1D")) +
    labs(
      title    = title_str,
      subtitle = sprintf(
        "Patient: age %f, AMH %.1f, AFC %f  |  90%% threshold shift: %+d eggs  |  95%% threshold shift: %+d eggs",
        eg_patient$age, eg_patient$amh, eg_patient$afc,
        as.integer(t90_new - t90_old),
        as.integer(t95_new - t95_old)
      ),
      x     = "Eggs retrieved",
      y     = "Density",
      fill  = "Model version",
      color = "Model version"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title    = element_text(face = "bold"),
      plot.subtitle = element_text(color = "grey40", size = 10),
      legend.position = "bottom"
    ) |> 
    suppressWarnings()
}
```

## Egg Model 1
```{r}
plot_comparison(mu1, theta_egg1, theta_old,
                "Egg 1-Cycle: Fitted vs. Original Theta")
```

## Egg Model 2
```{r}
plot_comparison(mu2, theta_egg2, theta_old,
                "Egg 2-Cycle: Fitted vs. Original Theta")
```

```{r resimulate_egg_model_1_draws}
# simulate_eggs <- function(newdata, theta_to_use, n = 2000) {
#   mu <- predict(egg_model, newdata, type = "response")
#   theta <- theta_to_use
#   rnbinom(n, mu = mu, size = theta)
# }
```

```{r compare_simulated_eggs_with_hardcoded_theta_and_model_theta_egg_model_1}
# model_theta_sims = simulate_eggs(df_no_na, egg_model$theta)
# hardcoded_theta_sims = simulate_eggs(df_no_na, 5)
# 
# compare_thresholds(model_theta_sims, theta_egg1, theta_old)

```

# Example Monte-Carlo Simulations
```{r}
new_patient <- eggmodel_sim_results[1,]
# previously_simulated_thresholds = df1_wit

#May have to change number of breaks depending on eggs, could adjust to 15 if fewer is needed
hist(simulate_eggs(new_patient), breaks = 30)
hist(simulate_eggs2(eggmodel2_sim_results[1,]), breaks = 30)

############################################################
# Histogram with Threshold lines (1-CYCLE)
############################################################

plot_egg_distribution <- function(newdata) {

  sims <- simulate_eggs(newdata)

  t90 <- quantile(sims, 0.10)
  t95 <- quantile(sims, 0.05)

  simulated_distribution = sims |> 
    table() |> 
    as.data.frame() |> 
    dplyr::rename(number_eggs_1_cycle = sims, count = Freq) |> 
    tidyr::as_tibble() |> 
    dplyr::mutate(the_simulated_eggs = number_eggs_1_cycle == newdata$sim_eggs_one_cycle)
  
  ggplot() + 
    geom_col(aes(x = number_eggs_1_cycle, y = count, fill = the_simulated_eggs), data = simulated_distribution) +
    geom_vline(xintercept = newdata$sim_egg_1cycle_90, lty = 2, col = 'orange', linewidth = 1) + 
    geom_vline(xintercept = newdata$sim_egg_1cycle_95, lty = 2, col = 'red', linewidth = 1) + 
    scale_fill_manual(values = c("TRUE" = "purple", "FALSE" = "darkgrey")) +
    labs(title = "Patient-level Simulated Number of Eggs Retrieved",
         subtitle = "Compared to simulated distribution based on age, AFC, and AMH (2000 MC runs)",
         caption = "Red line is 5th quantile and orange line is 10th quantile",
         x = "Simulated Egg Distribution in 1 Cycle",
         y = "Number of Simulated 'draws'") + 
    theme_minimal()
}

# Run it
plot_egg_distribution(new_patient)

```
