---
title: "Insurance Modeling Analysis"
author: "Charm Economics"
date: "`r Sys.Date()`"
output: 
  rmdformats::downcute
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(dplyr)
library(ggplot2)
library(VGAM)
library(pROC)
library(MASS)
library(ResourceSelection)
library(knitr)
library(tidyr)
library(readr)
set.seed(123)
```

```{r define_functions}
source(file.path(here::here(),"R/utils/modeling_stats_utility_functions.R"))
```

```{r read_in_and_clean_cycle_data, include=F}
df <- readr::read_rds("data/clean_data/cld_filtered.rds") |> 
  dplyr::rename(pregnant = pregnant_bool)

egg_df = readr::read_rds("data/clean_data/cleaned_egg_level_df.rds")

# Number of unique PAT ids
df = df |> dplyr::mutate(the_year = lubridate::year(treatment_start_date))
```

# Descriptive Statistics {.tabset}

## Number of Unique Patients

```{r}
df |> 
  dplyr::select(pat_id, age) |> 
  dplyr::distinct() |> 
  dplyr::count(age) |> 
  knitr::kable()
```

## Age, AMH and AFC Distributions
```{r}
suppressWarnings(
df |> 
  dplyr::select(age, afc, amh) |> 
  tidyr::pivot_longer(cols = dplyr::everything()) |> 
  dplyr::mutate(name = factor(name, levels = c('age','afc','amh'))) |> 
  ggplot(aes(value)) +
  facet_wrap(~ name, scales = 'free', ncol = 1) +
  geom_histogram(bins = 20) + 
  scale_x_continuous(breaks = scales::breaks_pretty())
)

df |> 
  dplyr::select(age, afc, amh) |> 
  tidyr::pivot_longer(cols = dplyr::everything()) |> 
  dplyr::group_by(name) |> 
  dplyr::reframe(maximum = max(value,na.rm=T), minimum = min(value, na.rm=T), median = median(value, na.rm=T)) |> 
  knitr::kable()
```


## Descriptive Pregnancy Rates
```{r}
pregnancy_rates <- df %>%
  group_by(age) %>%
  summarize(
    cycles = n(),
    pregnancies = sum(pregnant),
    pregnancy_rate = pregnancies / cycles
  )

kable(pregnancy_rates, digits = 3)

ggplot(pregnancy_rates, aes(x = age, y = pregnancy_rate)) +
  geom_col() +
  labs(
    title = "Pregnancy Rate by Age Group",
    x = "Age Group",
    y = "Pregnancy Rate"
  )
```

## Dataset Date Range
```{r}
df |> 
  dplyr::reframe(earliest_date = min(treatment_start_date, na.rm=T),
                 latest_date = max(treatment_start_date, na.rm=T)) |> 
  knitr::kable()

df |> 
  dplyr::mutate(year = lubridate::year(treatment_start_date)) |> 
  dplyr::count(year, name = 'unique patient/treatment cycle combinations')

ggplot(df) + 
  geom_histogram(aes(treatment_start_date)) + 
  scale_x_date(date_breaks = 'year', date_labels = '%Y') +
  theme_minimal()
```

## Key Variables
```{r make_tbl_for_key_variables}
var_summary <- df %>%
  mutate(
    total_2pn_eggs = num_2pn_from_fresh_eggs + num_2pn_from_frozen_eggs
  ) %>%
  group_by(age) %>%
  summarise(
    n_cycles = n(),
    
    # Euploid embryos
    mean_euploid    = mean(num_euploid_blast, na.rm = TRUE),
    iqr_euploid_l   = quantile(num_euploid_blast, 0.25, na.rm = TRUE),
    iqr_euploid_u   = quantile(num_euploid_blast, 0.75, na.rm = TRUE),
    pct_ge1_euploid = mean(coalesce(num_euploid_blast, 0) >= 1) * 100,
    
    # Eggs retrieved
    mean_eggs       = mean(num_eggs_collected, na.rm = TRUE),
    iqr_eggs_l      = quantile(num_eggs_collected, 0.25, na.rm = TRUE),
    iqr_eggs_u      = quantile(num_eggs_collected, 0.75, na.rm = TRUE),
    
    # MII eggs
    mean_m2         = mean(num_m2_eggs, na.rm = TRUE),
    iqr_m2_l        = quantile(num_m2_eggs, 0.25, na.rm = TRUE),
    iqr_m2_u        = quantile(num_m2_eggs, 0.75, na.rm = TRUE),
    
    # Fertilized (2PN)
    mean_2pn        = mean(total_2pn_eggs, na.rm = TRUE),
    iqr_2pn_l       = quantile(total_2pn_eggs, 0.25, na.rm = TRUE),
    iqr_2pn_u       = quantile(total_2pn_eggs, 0.75, na.rm = TRUE),
    
    # AMH
    mean_amh        = mean(amh, na.rm = TRUE),
    iqr_amh_l       = quantile(amh, 0.25, na.rm = TRUE),
    iqr_amh_u       = quantile(amh, 0.75, na.rm = TRUE),
    
    # AFC
    mean_afc        = mean(afc, na.rm = TRUE),
    iqr_afc_l       = quantile(afc, 0.25, na.rm = TRUE),
    iqr_afc_u       = quantile(afc, 0.75, na.rm = TRUE),
    
    .groups = "drop"
  )


var_summary_fmt <- var_summary %>%
  mutate(
    eggs_fmt = sprintf("%.1f (%.1f–%.1f)", mean_eggs, iqr_eggs_l, iqr_eggs_u),
    m2_fmt   = sprintf("%.1f (%.1f–%.1f)", mean_m2, iqr_m2_l, iqr_m2_u),
    p2n_fmt  = sprintf("%.1f (%.1f–%.1f)", mean_2pn, iqr_2pn_l, iqr_2pn_u),
    amh_fmt  = sprintf("%.2f (%.2f–%.2f)", mean_amh, iqr_amh_l, iqr_amh_u),
    afc_fmt  = sprintf("%.1f (%.1f–%.1f)", mean_afc, iqr_afc_l, iqr_afc_u),
    euploid_fmt = sprintf("%.1f (%.1f–%.1f)", mean_euploid, iqr_euploid_l, iqr_euploid_u),
    euploid_pct = sprintf("%.1f%%", pct_ge1_euploid)
  )

var_summary_fmt |> 
  knitr::kable()
```

# Removal of Patients over 43 years of age

```{r}
df = df |> 
  dplyr::filter(age <= 43)

egg_df = egg_df |> 
  dplyr::filter(age <= 43)
```

# Patient-level Statistical Models {.tabset}

Two models here:

Live birth model
Euploid embryo model


```{r make_model_1_dataset}

# <!-- This filter reduces the number of rows of data in this cycle-level table from `r nrow(df)` to `r nrow(m1df)`. -->
  
# Note - models in this section include only treatments of type 'ICSI' (the majority of the treatments) and NOT those of type 'IVF'.

euploid_embryos_from_egg_df = egg_df |> 
  dplyr::select(pat_id,tx_id,tx_cycle,pgd_clinresult) |> 
  dplyr::group_by(pat_id,tx_id) |> 
  dplyr::reframe(euploid_success = sum(pgd_clinresult %in% c("Euploid", "Normal") & !is.na(pgd_clinresult)) > 0)

m1df = df |>
  dplyr::left_join(euploid_embryos_from_egg_df) |> 
  dplyr::select(euploid_success, age, afc, amh, birth_bool) |> 
  tidyr::drop_na()

  # dplyr::filter(treatment == 'ICSI')
```



## Live-birth Model {.tabset}

### Model Outputs

```{r run_first_two_models}

# <!-- After dropping rows with NA for any of `euploid_success`, `age`, `afc`, or `amh`, the treatment cycle-level dataset has `r nrow(m1df)` observations, `r nrow(m1df_eup)` of which achieved a euploid embryo success. -->
  
euploid_model <- glm(
  euploid_success ~ age + amh + afc,
  family = binomial,
  data = m1df
)

live_birth_model <- glm(
  birth_bool ~ age + amh + afc,
  family = quasibinomial,
  data = m1df |> dplyr::filter(euploid_success)
)

m1df$pred_euploid <- predict(euploid_model, new_data = m1df, type = "response")

m1df_eup = m1df |> 
  dplyr::filter(euploid_success)

m1df_eup$pred_lb_given_euploid <- predict(
  live_birth_model,
  newdata = m1df_eup,
  type = "response"
)

m1df_eup$pred_live_birth <- m1df_eup$pred_euploid * m1df_eup$pred_lb_given_euploid

livebirth_sim_results <- m1df_eup |>
  dplyr::rowwise() |>
  dplyr::mutate(
    success_sim = rbinom(1, size = 1, prob = pred_live_birth),
    livebirth_success_simulated = success_sim >= 1
  ) |>
  dplyr::ungroup()
```

<!-- The live-birth model's measure of dispersion is `r check_dispersion(live_birth_model)`. -->

```{r}
livebirth_sim_results |>
  DT::datatable()

livebirth_sim_results |> 
  dplyr::count(livebirth_success_simulated) |> 
  dplyr::mutate(percentage = round( 100 * n / sum(n), 2)) |> 
  knitr::kable()
```

```{r read_in_egg_dataframe_and_clean_it, include = F}

# I'm unsure if we NEED to read in a new copy of the structured egg table... but if we do, we follow the same cleaning steps as above.


# Zoom in on relevant rows, drop rows with no info for age and attempt.
egg_df = egg_df |> 
  dplyr::select(pat_id, egg_id, age, attempt, pgd_clinresult, frozen_embryo, frozen_egg, afc_other, afc_first_baseline, amh, amh_start) |> 
  dplyr::filter(!is.na(age))

# Add treatment course ID and cycle number ID columns
df = df |> 
  dplyr::arrange(pat_id, treatment_start_date) |>
  dplyr::group_by(pat_id) |>
  dplyr::mutate(
    gap_days = as.numeric(treatment_start_date - dplyr::lag(treatment_start_date)),
    new_treat_course = ifelse(is.na(gap_days) | gap_days > 365, 1, 0),
    treat_course_id = cumsum(new_treat_course)
  ) |> 
  dplyr::select(-c(gap_days,new_treat_course)) |> 
  # Define cycle number within a treatment course
  dplyr::group_by(pat_id, treat_course_id) |>
  dplyr::arrange(treatment_start_date) |>
  # Cycle number is the key column we're making here - within a given 'treat_course', 
  # we can have various related cycles (cycle 1, 2, etc.)
  dplyr::mutate(cycle_number = dplyr::row_number()) |> 
  dplyr::ungroup()

cost_per_cycle <- 22014
payout_amount <- 17611
risk_margin <- 0.20
admin_load <- 0.10

n_sim <- 5000 # Monte Carlo iterations

```

### Simulated Probabilities averaged by Age
```{r}
livebirth_sim_results |> 
  dplyr::group_by(age) |> 
  dplyr::reframe(
    pct_livebirth_simulated = mean(livebirth_success_simulated, na.rm = TRUE)
  ) |> 
  dplyr::arrange(age) |> 
  knitr::kable()
```

## Euploid Embryo Model (Binomial) {.tabset}

### Model Outputs

<!-- The euploid model's measure of dispersion is `r check_dispersion(euploid_model)`. -->

```{r}
euploid_sim_results <- m1df |>
  dplyr::rowwise() |>
  dplyr::mutate(
    success_sim = rbinom(1, size = 1, prob = pred_euploid),
    ge1_euploid_sim = success_sim >= 1
  ) |>
  dplyr::ungroup()

euploid_sim_results |>
  DT::datatable()

euploid_sim_results |> 
  dplyr::count(ge1_euploid_sim) |> 
  dplyr::mutate(percentage = round( 100 * n / sum(n), 2)) |> 
  knitr::kable()
```

### Simulated Probabilities averaged by Age
```{r}
euploid_sim_results |> 
  dplyr::group_by(age) |> 
  dplyr::reframe(
    pct_1_plus_euploid_sim = mean(ge1_euploid_sim, na.rm = TRUE)
  ) |> 
  dplyr::arrange(age) |> 
  knitr::kable()
```

```{r}
# ### Plot of simulated payout
# payout_amount <- 22000
# 
# eup_df <- euploid_sim_results %>%
#   mutate(
#     pred_fail_p = 1 - pred_euploid,
#     pred_payout = pred_fail_p * payout_amount
#   ) %>%
#   group_by(age) %>%
#   summarise(
#     mean_pred_payout = mean(pred_payout, na.rm = TRUE),
#     sd_pred_payout = sd(pred_payout, na.rm = TRUE),
#     number_patients = n(),
#     .groups = "drop"
#   )
# 
# eup_df %>%
#   ggplot(aes(x = age, y = mean_pred_payout)) +
#   geom_smooth(linewidth = 1, color = "#2C7FB8") +
#   geom_point(size = 2.5, color = "#2C7FB8", alpha = 0.8) +
#   labs(
#     title = "Expected Insurance Payout by Age",
#     subtitle = "Euploid Embryo Model",
#     x = "Age (years)",
#     y = "Average Predicted Payout ($)"
#   ) +
#   scale_y_continuous(labels = scales::dollar_format()) +
#   scale_x_continuous(limits = c(30, 42), breaks = 30:42) +
#   theme_minimal(base_size = 13) +
#   theme(
#     plot.title = element_text(face = "bold", size = 16),
#     plot.subtitle = element_text(size = 12, color = "gray40"),
#     axis.title = element_text(face = "bold"),
#     axis.text = element_text(color = "gray20"),
#     panel.grid.minor = element_blank()
#   )
```

# Egg Actuarial Models (Beta-binomial / Binomial + Monte-Carlo Simulation) {.tabset}

## Single-cycle Egg Model (Binomial model) {.tabset}

```{r}
# Remove rows where amh is unknown.
df = df |> dplyr::filter(!is.na(amh))
```

### Model Outputs

Note: this model uses the patient / cycle-level table and does not currently filter away rows for eggs of cryo_day_2 or cryo_day_3; N of this dataframe is therefore `r nrow(df)`.

```{r}

df_no_na = df |> dplyr::select(pat_id, tx_id, num_eggs_collected, age, amh, afc) |> tidyr::drop_na()

# First, construct a Poisson-distribution of the number of eggs collected based on the data we have
egg_model <- glm(
  num_eggs_collected ~ age + amh + afc,
  family = quasipoisson(link = "log"),
  data = df_no_na
)

df_no_na$pred_eggmodel_1 = predict(egg_model, new_data = df_no_na, type = "response")

# Next, for each patient, based on their characteristics (age, afc, amh), simulate 
# 2000 hypothetical egg gathering treatments; these make up the simulated
# expectable distribution that is unique to that patient. We then take 
# out the 90th and 95th percentile of expectable eggs and use that as a 
# threshold for whether or not a simulated session would be successful
# for this patient. It's all about likelihoods!
egg_1cycle_simulated_thresholds_by_patient = df_no_na |> 
  dplyr::group_by(pat_id) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    sim_egg_1cycle_distribution = simulate_eggs(.x)
    sim_egg_1cycle_90 = quantile(sim_egg_1cycle_distribution, 0.1)
    sim_egg_1cycle_95 = quantile(sim_egg_1cycle_distribution, 0.05)
    output = data.frame(
      pat_id = .x$pat_id,
      sim_egg_1cycle_90,
      sim_egg_1cycle_95
    )
    row.names(output) = NULL
    output |> tidyr::as_tibble()
  },.progress) |> 
  dplyr::bind_rows() |> 
  dplyr::distinct() |> 
  suppressWarnings()

df_with_thresholds <- df_no_na |>
  dplyr::left_join(egg_1cycle_simulated_thresholds_by_patient, by = "pat_id")

# Now we simulate a single 'draw' for each patient, compare that draw to the 
# patient-specific distribution of expectable number of eggs, and classify
# the draw as either a success or a failure (i.e. above or below the threshold, respectively)
eggmodel_sim_results <- df_with_thresholds |>
  dplyr::rowwise() %>%
  dplyr::mutate(
    sim_eggs_one_cycle = simulate_eggs(., n = 1),
    success_90 = sim_eggs_one_cycle >= sim_egg_1cycle_90,
    success_95 = sim_eggs_one_cycle >= sim_egg_1cycle_95
  ) |>
  dplyr::ungroup()

eggmodel_sim_results |> 
  dplyr::count(success_90, success_95) |> 
  dplyr::mutate(percentage = round(100*n/sum(n),4))

```

### Simulated Probabilities averaged by Age
```{r}
eggmodel_sim_results |>
  dplyr::group_by(age) |>
  dplyr::reframe(
    pct_success_90 = mean(success_90, na.rm = TRUE),
    pct_success_95 = mean(success_95, na.rm = TRUE),
  ) |>
  dplyr::arrange(age) |>
  knitr::kable()
```

```{r}
### Plot of simulated payout
# payout_amount <- 22000
# 
# eup_df <- eggmodel_sim_results %>%
#   mutate(
#     pred_fail_p = 1 - pred_egg,
#     pred_payout = pred_fail_p * payout_amount
#   ) %>%
#   group_by(age) %>%
#   summarise(
#     mean_pred_payout = mean(pred_payout, na.rm = TRUE),
#     sd_pred_payout = sd(pred_payout, na.rm = TRUE),
#     number_patients = n(),
#     .groups = "drop"
#   )
# 
# eup_df %>%
#   ggplot(aes(x = age, y = mean_pred_payout)) +
#   geom_smooth(linewidth = 1, color = "#2C7FB8") +
#   geom_point(size = 2.5, color = "#2C7FB8", alpha = 0.8) +
#   labs(
#     title = "Expected Insurance Payout by Age",
#     subtitle = "Euploid Embryo Model",
#     x = "Age (years)",
#     y = "Average Predicted Payout ($)"
#   ) +
#   scale_y_continuous(labels = scales::dollar_format()) +
#   scale_x_continuous(limits = c(30, 42), breaks = 30:42) +
#   theme_minimal(base_size = 13) +
#   theme(
#     plot.title = element_text(face = "bold", size = 16),
#     plot.subtitle = element_text(size = 12, color = "gray40"),
#     axis.title = element_text(face = "bold"),
#     axis.text = element_text(color = "gray20"),
#     panel.grid.minor = element_blank()
#   )
```

<!-- The one-cycle egg model's measure of dispersion is `r check_dispersion(egg_model)`. -->

```{r}
# ### Plot Predicted payouts
# p3_df = single_cycle_egg_df |> 
#   dplyr::mutate(pred_fail_p = 1 - pred_p) |> 
#   dplyr::mutate(pred_payout = pred_fail_p * 20000) |>  # Arbitrary amount of payout!
#   dplyr::group_by(age, amh) |> 
#   dplyr::reframe(mean_pred_payout = mean(pred_payout, na.rm=T))
# 
# p3_df |> 
#   ggplot(aes(x = age, y = mean_pred_payout)) +
#   geom_line(aes(col = amh)) +
#   # geom_smooth(aes(group = amh_bin, colour = amh_bin), method = 'lm') +
#   geom_point(aes(colour = amh)) + 
#   # facet_wrap( ~ amh_bin, ncol = 1) +
#   labs(y = "Total Simulated Payout by Age", colour = 'AMH bin', x = "Age") +
#   scale_y_continuous(labels = scales::dollar) +
#   scale_x_continuous(limits = c(30,42)) + 
#   theme_minimal()
```

```{r}
### Simulated Probabilities averaged by Age
# s_egg_sim_results %>%
#   group_by(age) %>%
#   summarise(
#     prob_success = mean(success_sim, na.rm = TRUE),
#     .groups = "drop"
#   ) %>%
#   arrange(age) %>%
#   knitr::kable()
```

## Two-cycle Egg Model (Binomial model) {.tabset}

Note: this model includes only patients that have more than one treatment recorded in the dataset.
Note: this model assumes that treatments that occur within 365 days belong to a single treatment course, i.e. a collection of cycles including cycle 1, cycle 2, etc.; this assumption was necessary to compute this model because the dataset lacks a variable that otherwise links treatments across time. We could modify this assumption by removing this 365-day date proximity threshold, though this may incorrectly imply relatedness between distinct treatments performed by a patient across multiple years.

### Model Outputs

```{r}
df_2 <- df %>%
  dplyr::arrange(pat_id, treatment_start_date) %>%
  dplyr::group_by(pat_id) %>%
  dplyr::mutate(
    eggs_cycle1 = num_eggs_collected,
    eggs_cycle2 = dplyr::lead(num_eggs_collected),
    eggs_2cycle_total = eggs_cycle1 + eggs_cycle2
  ) %>%
  dplyr::filter(!is.na(eggs_2cycle_total)) %>%
  dplyr::ungroup()

# Ensure we don't have missing data for any of our variables.
df_2 = df_2 |> 
  dplyr::select(pat_id, eggs_2cycle_total, age, amh, afc, eggs_cycle1, eggs_cycle2, eggs_2cycle_total) |> 
  tidyr::drop_na()

egg2_model <- glm(
  eggs_2cycle_total ~ age + amh + afc + eggs_cycle1,
  family = quasipoisson(link = "log"),
  data = df_2
)

df_2$pred_eggmodel_2 = predict(egg2_model, new_data = df_2, type = "response")

# Next, for each patient, based on their characteristics (age, afc, amh), simulate 
# 2000 hypothetical egg gathering treatments; these make up the simulated
# expectable distribution that is unique to that patient. We then take 
# out the 90th and 95th percentile of expectable eggs and use that as a 
# threshold for whether or not a simulated session would be successful
# for this patient. It's all about likelihoods!
egg_2cycle_simulated_thresholds_by_patient = df_2 |> 
  dplyr::group_by(pat_id) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    sim_egg_2cycle_distribution = simulate_eggs2(.x)
    sim_egg_2cycle_90 = quantile(sim_egg_2cycle_distribution, 0.1)
    sim_egg_2cycle_95 = quantile(sim_egg_2cycle_distribution, 0.05)
    output = data.frame(
      pat_id = .x$pat_id,
      sim_egg_2cycle_90,
      sim_egg_2cycle_95
    )
    row.names(output) = NULL
    output |> tidyr::as_tibble()
  },.progress) |> 
  dplyr::bind_rows() |> 
  dplyr::distinct() |> 
  suppressWarnings()

df2_with_thresholds <- df_2 |>
  dplyr::left_join(egg_2cycle_simulated_thresholds_by_patient, by = "pat_id")

# Now we simulate a single 'draw' for each patient, compare that draw to the 
# patient-specific distribution of expectable number of eggs, and classify
# the draw as either a success or a failure (i.e. above or below the threshold, respectively)
eggmodel2_sim_results <- df2_with_thresholds |>
  dplyr::rowwise() %>%
  dplyr::mutate(
    sim_eggs_two_cycle = simulate_eggs2(., n = 1),
    success_90 = sim_eggs_two_cycle >= sim_egg_2cycle_90,
    success_95 = sim_eggs_two_cycle >= sim_egg_2cycle_95
  ) |>
  dplyr::ungroup()

eggmodel2_sim_results |> 
  dplyr::count(success_90, success_95) |> 
  dplyr::mutate(percentage = round(100*n/sum(n),4))

```

### Simulated Probabilities averaged by Age
```{r}
eggmodel2_sim_results |>
  dplyr::group_by(age) |>
  dplyr::reframe(
    pct_success_90 = mean(success_90, na.rm = TRUE),
    pct_success_95 = mean(success_95, na.rm = TRUE),
  ) |>
  dplyr::arrange(age) |>
  knitr::kable()
```

```{r, include = F}
# Euploid model ROC / AUC
roc_livebirth <- roc(livebirth_sim_results$livebirth_success_simulated, livebirth_sim_results$pred_live_birth)
auc_livebirth <- auc(roc_livebirth)
roc_euploid <- roc(m1df$euploid_success, m1df$pred_euploid)
auc_euploid <- auc(roc_euploid)
roc_egg1 <- roc(eggmodel_sim_results$success_90, eggmodel_sim_results$pred_eggmodel_1)
auc_egg1 <- auc(roc_egg1)
roc_egg2 <- roc(eggmodel2_sim_results$success_90, eggmodel2_sim_results$pred_eggmodel_2)
auc_egg2 <- auc(roc_egg2)

# --- INTERPRETATION GUIDE ---
# AUC < 0.60 → Poor model (not usable)
# AUC 0.60–0.70 → Weak (needs improvement or more predictors)
# AUC 0.70–0.80 → Acceptable
# AUC 0.80–0.90 → Strong
# AUC > 0.90 → Possibly overfit (check carefully)

# plot(roc_euploid)

# Dispersion checks
disp_livebirth = check_dispersion(live_birth_model)
disp_euploid <- check_dispersion(euploid_model)
disp_egg <- check_dispersion(egg_model)
disp_egg2 <- check_dispersion(egg2_model)

```

# Model Validation
```{r}
############################################################
# Validation Summary Table
############################################################

validation_summary <- data.frame(
  Model = c("Live Birth","Euploid", "Egg 1 Cycle", "Egg 2 Cycle"),
  AUC = c(auc_livebirth, auc_euploid, auc_egg1, auc_egg2),
  Dispersion = c(disp_livebirth, disp_euploid, disp_egg, disp_egg2)
)

validation_summary |> knitr::kable()
```

# Example Monte-Carlo Simulations
```{r}
new_patient <- eggmodel_sim_results[1,]
# previously_simulated_thresholds = df1_wit

#May have to change number of breaks depending on eggs, could adjust to 15 if fewer is needed
hist(simulate_eggs(new_patient), breaks = 30)
hist(simulate_eggs2(eggmodel2_sim_results[1,]), breaks = 30)

############################################################
# Histogram with Threshold lines (1-CYCLE)
############################################################

plot_egg_distribution <- function(newdata) {

  sims <- simulate_eggs(newdata)

  t90 <- quantile(sims, 0.10)
  t95 <- quantile(sims, 0.05)

  simulated_distribution = sims |> 
    table() |> 
    as.data.frame() |> 
    dplyr::rename(number_eggs_1_cycle = sims, count = Freq) |> 
    tidyr::as_tibble() |> 
    dplyr::mutate(the_simulated_eggs = number_eggs_1_cycle == newdata$sim_eggs_one_cycle)
  
  ggplot() + 
    geom_col(aes(x = number_eggs_1_cycle, y = count, fill = the_simulated_eggs), data = simulated_distribution) +
    geom_vline(xintercept = newdata$sim_egg_1cycle_90, lty = 2, col = 'orange', linewidth = 1) + 
    geom_vline(xintercept = newdata$sim_egg_1cycle_95, lty = 2, col = 'red', linewidth = 1) + 
    scale_fill_manual(values = c("TRUE" = "purple", "FALSE" = "darkgrey")) +
    labs(title = "Patient-level Simulated Number of Eggs Retrieved",
         subtitle = "Compared to simulated distribution based on age, AFC, and AMH (2000 MC runs)",
         caption = "Red line is 5th quantile and orange line is 10th quantile",
         x = "Simulated Egg Distribution in 1 Cycle",
         y = "Number of Simulated 'draws'") + 
    theme_minimal()
}

# Run it
plot_egg_distribution(new_patient)

```

```{r}
# ### Plot Predicted payouts
# 
# p4_df = df_multiple |> 
#   dplyr::mutate(pred_fail_p = 1 - pred_p) |> 
#   dplyr::mutate(pred_payout = pred_fail_p * 20000) |>  # Arbitrary amount of payout!
#   dplyr::group_by(age, amh) |> 
#   dplyr::reframe(mean_pred_payout = mean(pred_payout, na.rm=T))
# 
# p4_df |> 
#   ggplot(aes(x = age, y = mean_pred_payout)) +
#   geom_line(aes(col = amh)) +
#   # geom_smooth(aes(group = amh_bin, colour = amh_bin), method = 'lm') +
#   geom_point(aes(colour = amh)) + 
#   # facet_wrap( ~ amh_bin, ncol = 1) +
#   labs(y = "Total Simulated Payout by Age", colour = 'AMH bin', x = "Age") +
#   scale_y_continuous(labels = scales::dollar) +
#   scale_x_continuous(limits = c(30,42)) + 
#   theme_minimal()
```

```{r}
# ### Simulated Probabilities averaged by Age
# 
# df_multiple_sim_results %>%
#   group_by(age) %>%
#   summarise(
#     prob_fifteen_eggs_or_more_sim = mean(success_sim, na.rm = TRUE),
#     .groups = "drop"
#   ) %>%
#   arrange(age) %>%
#   knitr::kable()
```

```{r}
# # Pricing Outputs
# calc_pricing <- function(expected_payout_vec, model_name) {
#   tibble::tibble(
#     model = model_name,
#     expected_cost = mean(expected_payout_vec, na.rm = TRUE),
#     var_95 = quantile(expected_payout_vec, 0.95, na.rm = TRUE),
#     premium = expected_cost * (1 + risk_margin + admin_load)
#   )
# }
# 
# # Use the same payout amount consistently
# payout_amount <- 22000
# 
# # 1. Clinical pregnancy model
# live_birth_costs <- live_birth_sim_results %>%
#   mutate(
#     pred_fail_p = 1 - pred_p,
#     expected_payout = pred_fail_p * payout_amount
#   )
# 
# live_birth_results <- calc_pricing(
#   expected_payout_vec = live_birth_costs$expected_payout,
#   model_name = "Live-birth Model"
# )
# 
# # 2. Euploid embryo model
# euploid_costs <- euploid_sim_results %>%
#   mutate(
#     pred_fail_p = 1 - prob_ge1_euploid,
#     expected_payout = pred_fail_p * payout_amount
#   )
# 
# euploid_results <- calc_pricing(
#   expected_payout_vec = euploid_costs$expected_payout,
#   model_name = "Euploid Embryo"
# )
# 
# # 3. Single-cycle egg model
# s_egg_costs <- single_cycle_egg_df %>%
#   mutate(
#     pred_fail_p = 1 - pred_p,
#     expected_payout = pred_fail_p * payout_amount
#   )
# 
# egg_results <- calc_pricing(
#   expected_payout_vec = s_egg_costs$expected_payout,
#   model_name = "Single-Cycle Egg"
# )
# 
# # 4. Two-cycle egg model
# two_egg_costs <- df_multiple %>%
#   mutate(
#     pred_fail_p = 1 - pred_p,
#     expected_payout = pred_fail_p * payout_amount
#   )
# 
# two_results <- calc_pricing(
#   expected_payout_vec = two_egg_costs$expected_payout,
#   model_name = "Two-Cycle Egg"
# )
```

```{r}
# # Final Comparison
# final_results <- bind_rows(
#   live_birth_results,
#   euploid_results,
#   egg_results,
#   two_results
# ) %>%
#   mutate(
#     expected_cost = scales::dollar(expected_cost),
#     var_95 = scales::dollar(var_95),
#     premium = scales::dollar(premium)
#   )
# 
# knitr::kable(final_results)
```
