---
title: "Insurance Modeling Analysis"
author: "Charm Economics"
date: "`r Sys.Date()`"
output: 
  rmdformats::downcute
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(dplyr)
library(ggplot2)
library(VGAM)
library(pROC)
library(MASS)
library(ResourceSelection)
library(knitr)
library(tidyr)
library(readr)
library(rprojroot)

knitr::opts_knit$set(
  root.dir = rprojroot::find_rstudio_root_file())
  
set.seed(123)

# Must match age_floor in PipelineA_01. Under-32 patients are treated as 32.
age_floor <- 32
```

```{r define_functions}
source(file.path(here::here(),"R/utils/modeling_stats_utility_functions.R"))
source(file.path(here::here(),"R/utils/model_recalibration.R"))
```

```{r read_in_and_clean_cycle_data, include=F}
df <- readr::read_rds("data/clean_data/cld_filtered.rds") |> 
  dplyr::rename(pregnant = pregnant_bool)

egg_df = readr::read_rds("data/clean_data/cleaned_egg_level_df.rds")

# Number of unique PAT ids
df = df |> dplyr::mutate(the_year = lubridate::year(treatment_start_date))
```

# Sections {.tabset}

## Descriptive Statistics {.tabset}

### Number of Unique Patients

```{r}
df |> 
  dplyr::select(pat_id, age) |> 
  dplyr::distinct() |> 
  dplyr::count(age) |> 
  knitr::kable()
```

### Age, AMH and AFC Distributions
```{r}
suppressWarnings(
df |> 
  dplyr::select(age, afc, amh) |> 
  tidyr::pivot_longer(cols = dplyr::everything()) |> 
  dplyr::mutate(name = factor(name, levels = c('age','afc','amh'))) |> 
  ggplot(aes(value)) +
  facet_wrap(~ name, scales = 'free', ncol = 1) +
  geom_histogram(bins = 20) + 
  scale_x_continuous(breaks = scales::breaks_pretty())
)

df |> 
  dplyr::select(age, afc, amh) |> 
  tidyr::pivot_longer(cols = dplyr::everything()) |> 
  dplyr::group_by(name) |> 
  dplyr::reframe(maximum = max(value,na.rm=T), minimum = min(value, na.rm=T), median = median(value, na.rm=T)) |> 
  knitr::kable()
```


### Descriptive Pregnancy Rates
```{r}
pregnancy_rates <- df %>%
  group_by(age) %>%
  summarize(
    cycles = n(),
    pregnancies = sum(pregnant),
    pregnancy_rate = pregnancies / cycles
  )

kable(pregnancy_rates, digits = 3)

ggplot(pregnancy_rates, aes(x = age, y = pregnancy_rate)) +
  geom_col() +
  labs(
    title = "Pregnancy Rate by Age Group",
    x = "Age Group",
    y = "Pregnancy Rate"
  )
```

### Dataset Date Range
```{r}
df |> 
  dplyr::reframe(earliest_date = min(treatment_start_date, na.rm=T),
                 latest_date = max(treatment_start_date, na.rm=T)) |> 
  knitr::kable()

df |> 
  dplyr::mutate(year = lubridate::year(treatment_start_date)) |> 
  dplyr::count(year, name = 'unique patient/treatment cycle combinations')

ggplot(df) + 
  geom_histogram(aes(treatment_start_date)) + 
  scale_x_date(date_breaks = 'year', date_labels = '%Y') +
  theme_minimal()
```

### Key Variables
```{r make_tbl_for_key_variables}
var_summary <- df %>%
  mutate(
    total_2pn_eggs = num_2pn_from_fresh_eggs + num_2pn_from_frozen_eggs
  ) %>%
  group_by(age) %>%
  summarise(
    n_cycles = n(),
    
    # Euploid embryos
    mean_euploid    = mean(num_euploid_blast, na.rm = TRUE),
    iqr_euploid_l   = quantile(num_euploid_blast, 0.25, na.rm = TRUE),
    iqr_euploid_u   = quantile(num_euploid_blast, 0.75, na.rm = TRUE),
    pct_ge1_euploid = mean(coalesce(num_euploid_blast, 0) >= 1) * 100,
    
    # Eggs retrieved
    mean_eggs       = mean(num_eggs_collected, na.rm = TRUE),
    iqr_eggs_l      = quantile(num_eggs_collected, 0.25, na.rm = TRUE),
    iqr_eggs_u      = quantile(num_eggs_collected, 0.75, na.rm = TRUE),
    
    # MII eggs
    mean_m2         = mean(num_m2_eggs, na.rm = TRUE),
    iqr_m2_l        = quantile(num_m2_eggs, 0.25, na.rm = TRUE),
    iqr_m2_u        = quantile(num_m2_eggs, 0.75, na.rm = TRUE),
    
    # Fertilized (2PN)
    mean_2pn        = mean(total_2pn_eggs, na.rm = TRUE),
    iqr_2pn_l       = quantile(total_2pn_eggs, 0.25, na.rm = TRUE),
    iqr_2pn_u       = quantile(total_2pn_eggs, 0.75, na.rm = TRUE),
    
    # AMH
    mean_amh        = mean(amh, na.rm = TRUE),
    iqr_amh_l       = quantile(amh, 0.25, na.rm = TRUE),
    iqr_amh_u       = quantile(amh, 0.75, na.rm = TRUE),
    
    # AFC
    mean_afc        = mean(afc, na.rm = TRUE),
    iqr_afc_l       = quantile(afc, 0.25, na.rm = TRUE),
    iqr_afc_u       = quantile(afc, 0.75, na.rm = TRUE),
    
    .groups = "drop"
  )


var_summary_fmt <- var_summary %>%
  mutate(
    eggs_fmt = sprintf("%.1f (%.1f–%.1f)", mean_eggs, iqr_eggs_l, iqr_eggs_u),
    m2_fmt   = sprintf("%.1f (%.1f–%.1f)", mean_m2, iqr_m2_l, iqr_m2_u),
    p2n_fmt  = sprintf("%.1f (%.1f–%.1f)", mean_2pn, iqr_2pn_l, iqr_2pn_u),
    amh_fmt  = sprintf("%.2f (%.2f–%.2f)", mean_amh, iqr_amh_l, iqr_amh_u),
    afc_fmt  = sprintf("%.1f (%.1f–%.1f)", mean_afc, iqr_afc_l, iqr_afc_u),
    euploid_fmt = sprintf("%.1f (%.1f–%.1f)", mean_euploid, iqr_euploid_l, iqr_euploid_u),
    euploid_pct = sprintf("%.1f%%", pct_ge1_euploid)
  )

var_summary_fmt |> 
  knitr::kable()
```

## Removal of Patients over 43 years of age

Just a flag to remind us that we have done this!

```{r}
df = df |> 
  dplyr::filter(age <= 43)

egg_df = egg_df |> 
  dplyr::filter(age <= 43)
```

## Patient-level Statistical Models {.tabset}

Two models here:

Live birth model
Euploid embryo model

```{r make_model_1_dataset}

# Note - models in this section include only treatments of type 'ICSI' (the majority of the treatments) and NOT those of type 'IVF'.

euploid_embryos_from_egg_df = egg_df |> 
  dplyr::select(pat_id,tx_id,tx_cycle,pgd_clinresult) |> 
  dplyr::group_by(pat_id,tx_id) |> 
  dplyr::reframe(euploid_success = sum(pgd_clinresult %in% c("Euploid", "Normal") & !is.na(pgd_clinresult)) > 0)

m1df = df |>
  dplyr::left_join(euploid_embryos_from_egg_df) |> 
  dplyr::arrange(pat_id, treatment_start_date) |> 
  dplyr::group_by(pat_id) |> 
  dplyr::mutate(cycle_number = dplyr::row_number()) |> 
  dplyr::ungroup() |> 
  dplyr::select(pat_id, euploid_success, age, afc, amh, birth_bool, cycle_number, treatment_start_date) |> 
  tidyr::drop_na() |> 
  # -- consistency with PipelineA_01: same eligibility exclusion + age floor,
  #    so this report describes the same population that is priced/validated.
  dplyr::filter(amh <= 12 & afc <= 40) |> 
  dplyr::mutate(age_model = pmax(age, age_floor))

```

### Live-birth Model {.tabset}

#### Model Outputs

```{r run_first_two_models}

# <!-- After dropping rows with NA for any of `euploid_success`, `age`, `afc`, or `amh`, the treatment cycle-level dataset has `r nrow(m1df)` observations, `r nrow(m1df_eup)` of which achieved a euploid embryo success. -->
  
# -- consistency: use the floored deployed models saved by PipelineA_01 --
# Previously these were refit here on RAW age (and on a different data slice),
# which is why the report disagreed with the priced/validated models. They are
# now loaded so everything downstream uses the exact same fitted objects.
#   euploid_model    : euploid_success ~ age_model + amh + afc + amh:afc  (binomial)
#   live_birth_model : birth_bool      ~ age_model * amh * afc            (binomial)
euploid_model    <- readr::read_rds("data/models/euploid_model_revised_modeling_statistics_end.rds")
live_birth_model <- readr::read_rds("data/models/livebirth_model_revised_modeling_statistics_end.rds")

# ############################################################
# # Export model coefficients for Excel pricing tool
# ############################################################
# 
# euploid_coef_export <- coef(euploid_model) |>
#   as.data.frame() |>
#   tibble::rownames_to_column("term") |>
#   dplyr::rename(estimate = 2) |>
#   dplyr::mutate(model = "euploid")
# 
# readr::write_csv(
#   euploid_coef_export,
#   "outputs/euploid_model_coefficients.csv"
# )
# 
# live_birth_coef_export <- coef(live_birth_model) |>
#   as.data.frame() |>
#   tibble::rownames_to_column("term") |>
#   dplyr::rename(estimate = 2) |>
#   dplyr::mutate(model = "live_birth_given_euploid")
# 
# readr::write_csv(
#   live_birth_coef_export,
#   "outputs/live_birth_model_coefficients.csv"
# )

m1df$pred_euploid <- predict(euploid_model, newdata = m1df, type = "response")
# m1df$pred_euploid <- predict(euploid_model, new_data = m1df, type = "response")

# Before applying recalibration, get metrics for model.
euploid_pre_calib_metrics = get_metrics(euploid_model, m1df, 'euploid_success')

# Apply recalibration corrections to these predicted values.
euploid_cal_model_fit_res = fit_and_apply_calibration_model(m1df$euploid_success, m1df$pred_euploid)
m1df$pred_euploid_recal = euploid_cal_model_fit_res$preds |> as.numeric()
euploid_calib_interc = 0.015
euploid_calib_coef = 0.853

# After applying recalibration, get metrics for model.
euploid_post_calib_metrics = get_metrics(euploid_cal_model_fit_res$model, m1df, 'euploid_success')

m1df_eup = m1df |> 
  dplyr::filter(euploid_success)

m1df_eup$pred_lb_given_euploid <- predict(
  live_birth_model,
  newdata = m1df_eup,
  type = "response"
)

# Before applying recalibration, get metrics for model.
lb_pre_calib_metrics = get_metrics(live_birth_model, m1df_eup, 'birth_bool')

live_birth_cal_model_fit_res = fit_and_apply_calibration_model(m1df_eup$birth_bool, m1df_eup$pred_lb_given_euploid)
m1df_eup$pred_lb_given_euploid_recal = live_birth_cal_model_fit_res$preds
lb_calib_interc = live_birth_cal_model_fit_res$intercept
lb_calib_coef = live_birth_cal_model_fit_res$slope

# Before applying recalibration, get metrics for model.
lb_post_calib_metrics = get_metrics(live_birth_cal_model_fit_res$model, m1df_eup, 'birth_bool')

m1df_eup$pred_live_birth_recal <- m1df_eup$pred_euploid_recal * m1df_eup$pred_lb_given_euploid_recal

livebirth_sim_results <- m1df_eup |>
  dplyr::rowwise() |>
  dplyr::mutate(
    success_sim = rbinom(1, size = 1, prob = pred_live_birth_recal),
    livebirth_success_simulated = success_sim >= 1
  ) |>
  dplyr::ungroup()
```

<!-- The live-birth model's measure of dispersion is `r check_dispersion(live_birth_model)`. -->

```{r}
livebirth_sim_results |>
  DT::datatable()

livebirth_sim_results |> 
  dplyr::count(livebirth_success_simulated) |> 
  dplyr::mutate(percentage = round( 100 * n / sum(n), 2)) |> 
  knitr::kable()
```

```{r read_in_egg_dataframe_and_clean_it, include = F}

# I'm unsure if we NEED to read in a new copy of the structured egg table... but if we do, we follow the same cleaning steps as above.


# Zoom in on relevant rows, drop rows with no info for age and attempt.
egg_df = egg_df |> 
  dplyr::select(pat_id, egg_id, age, attempt, pgd_clinresult, frozen_embryo, frozen_egg, afc_other, afc_first_baseline, amh, amh_start) |> 
  dplyr::filter(!is.na(age))

# Add treatment course ID and cycle number ID columns
df = df |> 
  dplyr::arrange(pat_id, treatment_start_date) |>
  dplyr::group_by(pat_id) |>
  dplyr::mutate(
    gap_days = as.numeric(treatment_start_date - dplyr::lag(treatment_start_date)),
    new_treat_course = ifelse(is.na(gap_days) | gap_days > 365, 1, 0),
    treat_course_id = cumsum(new_treat_course)
  ) |> 
  dplyr::select(-c(gap_days,new_treat_course)) |> 
  # Define cycle number within a treatment course
  dplyr::group_by(pat_id, treat_course_id) |>
  dplyr::arrange(treatment_start_date) |>
  # Cycle number is the key column we're making here - within a given 'treat_course', 
  # we can have various related cycles (cycle 1, 2, etc.)
  dplyr::mutate(cycle_number = dplyr::row_number()) |> 
  dplyr::ungroup()

cost_per_cycle <- 22014
payout_amount <- 17611
risk_margin <- 0.20
admin_load <- 0.10

n_sim <- 5000 # Monte Carlo iterations

```

#### Multi-cycle (Cumulative) Probabilities
```{r}
# --- Closed-form cumulative probability (Live Birth) ---
livebirth_multi_cycle = m1df_eup |> 
 dplyr::group_by(pat_id) |>
  dplyr::summarise(
    p1 = first(pred_live_birth_recal)
  ) |>
  dplyr::mutate(
    one_cycle_scs = 1 - (1 - p1)^1,
    two_cycles_scs = 1 - (1 - p1)^2,
    three_cycles_scs = 1 - (1 - p1)^3,
    one_cycle_fail = (1 - p1)^1,
    two_cycles_fail = (1 - p1)^2,
    three_cycles_fail = (1 - p1)^3
  )

livebirth_multi_cycle |> 
  DT::datatable()

readr::write_csv(livebirth_multi_cycle, "outputs/livebirth_multicycle_probabilities.csv")

```

#### Simulated Probabilities averaged by Age
```{r}
livebirth_sim_results |> 
  dplyr::group_by(age) |> 
  dplyr::reframe(
    pct_livebirth_simulated = mean(livebirth_success_simulated, na.rm = TRUE)
  ) |> 
  dplyr::arrange(age) |> 
  knitr::kable()
```

### Euploid Embryo Model (Binomial) {.tabset}

#### Model Outputs

<!-- The euploid model's measure of dispersion is `r check_dispersion(euploid_model)`. -->

```{r}
euploid_sim_results <- m1df |>
  dplyr::rowwise() |>
  dplyr::mutate(
    success_sim = rbinom(1, size = 1, prob = pred_euploid_recal),
    ge1_euploid_sim = success_sim >= 1
  ) |>
  dplyr::ungroup()

euploid_sim_results |>
  DT::datatable()

euploid_sim_results |> 
  dplyr::count(ge1_euploid_sim) |> 
  dplyr::mutate(percentage = round( 100 * n / sum(n), 2)) |> 
  knitr::kable()
```

#### Multi-cycle (Cumulative) Probabilities
```{r}
# --- Closed-form cumulative probability (Live Birth) ---
euploid_multi_cycle = m1df |> 
 dplyr::group_by(pat_id) |>
  dplyr::summarise(
    p1 = first(pred_euploid_recal)
  ) |>
  dplyr::mutate(
    one_cycle_scs = 1 - (1 - p1)^1,
    two_cycles_scs = 1 - (1 - p1)^2,
    three_cycles_scs = 1 - (1 - p1)^3,
    one_cycle_fail = (1 - p1)^1,
    two_cycles_fail = (1 - p1)^2,
    three_cycles_fail = (1 - p1)^3
  )

euploid_multi_cycle |> 
  DT::datatable()

readr::write_csv(euploid_multi_cycle, "outputs/euploid_embryo_multicycle_probabilities.csv")

```

#### Simulated Probabilities averaged by Age
```{r}
euploid_sim_results |> 
  dplyr::group_by(age) |> 
  dplyr::reframe(
    pct_1_plus_euploid_sim = mean(ge1_euploid_sim, na.rm = TRUE)
  ) |> 
  dplyr::arrange(age) |> 
  knitr::kable()
```

### Pre and Post Calibration Model Metrics
```{r}
calib_results = data.frame(
  model = c(rep('euploid',2), rep('livebirth',2)),
  pre_or_post = rep(c('pre','post'),2)
)
tbl_from_results_vec = function(vec){
  # browser()
  o = vec |> 
    as.data.frame()
  o$var_name = row.names(o)
  o |> 
    tidyr::pivot_wider(names_from = var_name, values_from = vec)
}

the_metrics = list(euploid_pre_calib_metrics, 
     euploid_post_calib_metrics,
     lb_pre_calib_metrics,
     lb_post_calib_metrics) |> 
  purrr::map(
    \(x) tbl_from_results_vec(x)
  ) |> 
  dplyr::bind_rows()

calib_results = calib_results |> 
  dplyr::bind_cols(the_metrics)

calib_results |> 
  knitr::kable()

# Export all the necessary bits for the Excel calculator.
# Coefficients (for p_raw) - compute p_raw 
# Calibration intercept (α)
# Calibration slope (β)
# Apply recalibration function

euploid_coef_export <- coef(euploid_model) |>
  as.data.frame() |>
  tibble::rownames_to_column("term") |>
  dplyr::rename(estimate = 2) |>
  dplyr::mutate(model = "euploid")

euploid_coef_export = euploid_coef_export |> 
  dplyr::bind_rows(
    calib_results |> dplyr::filter(model == 'euploid')
  )

# -- removed: PipelineA_03 (master) is the single owner of the coefficient CSVs,
#    and it now exports the SAME floored coefficients. Kept object for the report.
# readr::write_csv(euploid_coef_export, "outputs/euploid_model_coefficients.csv")

live_birth_coef_export <- coef(live_birth_model) |>
  as.data.frame() |>
  tibble::rownames_to_column("term") |>
  dplyr::rename(estimate = 2) |>
  dplyr::mutate(model = "live_birth_given_euploid")

live_birth_coef_export = live_birth_coef_export |> 
  dplyr::bind_rows(
    calib_results |> dplyr::filter(model == 'livebirth')
  )

# -- removed: PipelineA_03 (master) is the single owner of the coefficient CSVs.
# readr::write_csv(live_birth_coef_export, "outputs/live_birth_model_coefficients.csv")
```

## Egg Actuarial Models (Beta-binomial / Binomial + Monte-Carlo Simulation) {.tabset}

### Single-cycle Egg Model (Binomial model) {.tabset}

```{r}
# Remove rows where amh is unknown.
df = df |> dplyr::filter(!is.na(amh))
```

#### Model Outputs

Note: this model uses the patient / cycle-level table and does not currently filter away rows for eggs of cryo_day_2 or cryo_day_3; N of this dataframe is therefore `r nrow(df)`.

```{r}

df_no_na = df |> dplyr::select(pat_id, tx_id, num_eggs_collected, age, amh, afc) |> tidyr::drop_na() |> 
  # -- consistency with PipelineA_01: eligibility exclusion, 99% egg-count cap, age floor --
  dplyr::filter(amh <= 12 & afc <= 40) |> 
  dplyr::filter(num_eggs_collected <= round(quantile(num_eggs_collected, 0.99))) |> 
  dplyr::mutate(age_model = pmax(age, age_floor))

# -- consistency: load the floored egg-1 model saved by PipelineA_01 --
#   egg_model : num_eggs_collected ~ age_model + amh + afc + age_model:afc  (neg. binomial)
egg_model <- readr::read_rds("data/models/egg_1_model_revised_modeling_statistics_end.rds")

df_no_na$pred_eggmodel_1 = predict(egg_model, newdata = df_no_na, type = "response")

# Next, for each patient, based on their characteristics (age, afc, amh), simulate 
# 2000 hypothetical egg gathering treatments; these make up the simulated
# expectable distribution that is unique to that patient. We then take 
# out the 90th and 95th percentile of expectable eggs and use that as a 
# threshold for whether or not a simulated session would be successful
# for this patient. It's all about likelihoods!
egg_1cycle_simulated_thresholds_by_patient = df_no_na |> 
  dplyr::group_by(pat_id) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    sim_egg_1cycle_distribution = simulate_eggs(.x)
    sim_egg_1cycle_90 = quantile(sim_egg_1cycle_distribution, 0.1)
    sim_egg_1cycle_95 = quantile(sim_egg_1cycle_distribution, 0.05)
    output = data.frame(
      pat_id = .x$pat_id,
      sim_egg_1cycle_90,
      sim_egg_1cycle_95
    )
    row.names(output) = NULL
    output |> tidyr::as_tibble()
  },.progress) |> 
  dplyr::bind_rows() |> 
  dplyr::distinct() |> 
  suppressWarnings()

df_with_thresholds <- df_no_na |>
  dplyr::left_join(egg_1cycle_simulated_thresholds_by_patient, by = "pat_id")

# Now we simulate a single 'draw' for each patient, compare that draw to the 
# patient-specific distribution of expectable number of eggs, and classify
# the draw as either a success or a failure (i.e. above or below the threshold, respectively)
eggmodel_sim_results <- df_with_thresholds |>
  dplyr::rowwise() %>%
  dplyr::mutate(
    sim_eggs_one_cycle = simulate_eggs(., n = 1),
    success_90 = sim_eggs_one_cycle >= sim_egg_1cycle_90,
    success_95 = sim_eggs_one_cycle >= sim_egg_1cycle_95
  ) |>
  dplyr::ungroup()

eggmodel_sim_results |> 
  dplyr::count(success_90, success_95) |> 
  dplyr::mutate(percentage = round(100*n/sum(n),4))

# Conversion of simulated successes/failures to predicted probabilities so we can
# apply the same model tests and validations to the egg models as we do to 
# the live-birth and euploid embryo eggs.
eggmodel_sim_results <- eggmodel_sim_results %>%
  dplyr::mutate(
    pred_prob_success_90 = estimate_prob_success(
      ., simulate_eggs, "sim_egg_1cycle_90", n_sim = 1000
    ),
    pred_prob_success_95 = estimate_prob_success(
      ., simulate_eggs, "sim_egg_1cycle_95", n_sim = 1000
    )
  )
```

#### Simulated Probabilities averaged by Age
```{r}
eggmodel_sim_results |>
  dplyr::group_by(age) |>
  dplyr::reframe(
    pct_success_90 = mean(success_90, na.rm = TRUE),
    pct_success_95 = mean(success_95, na.rm = TRUE),
  ) |>
  dplyr::arrange(age) |>
  knitr::kable()
```

### Two-cycle Egg Model (Binomial model) {.tabset}

Note: this model includes only patients that have more than one treatment recorded in the dataset.
Note: this model assumes that treatments that occur within 365 days belong to a single treatment course, i.e. a collection of cycles including cycle 1, cycle 2, etc.; this assumption was necessary to compute this model because the dataset lacks a variable that otherwise links treatments across time. We could modify this assumption by removing this 365-day date proximity threshold, though this may incorrectly imply relatedness between distinct treatments performed by a patient across multiple years.

#### Model Outputs

```{r}
df_2 <- df %>%
  dplyr::arrange(pat_id, treatment_start_date) %>%
  dplyr::group_by(pat_id) %>%
  dplyr::mutate(
    eggs_cycle1 = num_eggs_collected,
    eggs_cycle2 = dplyr::lead(num_eggs_collected),
    eggs_2cycle_total = eggs_cycle1 + eggs_cycle2
  ) %>%
  dplyr::filter(!is.na(eggs_2cycle_total)) %>%
  dplyr::ungroup()

# Ensure we don't have missing data for any of our variables.
df_2 = df_2 |> 
  dplyr::select(pat_id, eggs_2cycle_total, age, amh, afc, eggs_cycle1, eggs_cycle2, eggs_2cycle_total) |> 
  tidyr::drop_na() |> 
  # -- consistency with PipelineA_01: eligibility exclusion, 99% 2-cycle cap, age floor --
  dplyr::filter(amh <= 12 & afc <= 40) |> 
  dplyr::filter(eggs_2cycle_total <= round(quantile(eggs_2cycle_total, 0.99))) |> 
  dplyr::mutate(age_model = pmax(age, age_floor))

# -- consistency: load the floored egg-2 model saved by PipelineA_01 --
#   egg2_model : eggs_2cycle_total ~ age_model + amh + afc + eggs_cycle1 + amh:afc  (neg. binomial)
egg2_model <- readr::read_rds("data/models/egg_2_model_revised_modeling_statistics_end.rds")

df_2$pred_eggmodel_2 = predict(egg2_model, newdata = df_2, type = "response")

# Next, for each patient, based on their characteristics (age, afc, amh), simulate 
# 2000 hypothetical egg gathering treatments; these make up the simulated
# expectable distribution that is unique to that patient. We then take 
# out the 90th and 95th percentile of expectable eggs and use that as a 
# threshold for whether or not a simulated session would be successful
# for this patient. It's all about likelihoods!
egg_2cycle_simulated_thresholds_by_patient = df_2 |> 
  dplyr::group_by(pat_id) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    sim_egg_2cycle_distribution = simulate_eggs2(.x)
    sim_egg_2cycle_90 = quantile(sim_egg_2cycle_distribution, 0.1)
    sim_egg_2cycle_95 = quantile(sim_egg_2cycle_distribution, 0.05)
    output = data.frame(
      pat_id = .x$pat_id,
      sim_egg_2cycle_90,
      sim_egg_2cycle_95
    )
    row.names(output) = NULL
    output |> tidyr::as_tibble()
  },.progress) |> 
  dplyr::bind_rows() |> 
  dplyr::distinct() |> 
  suppressWarnings()

df2_with_thresholds <- df_2 |>
  dplyr::left_join(egg_2cycle_simulated_thresholds_by_patient, by = "pat_id")

# Now we simulate a single 'draw' for each patient, compare that draw to the 
# patient-specific distribution of expectable number of eggs, and classify
# the draw as either a success or a failure (i.e. above or below the threshold, respectively)
eggmodel2_sim_results <- df2_with_thresholds |>
  dplyr::rowwise() %>%
  dplyr::mutate(
    sim_eggs_two_cycle = simulate_eggs2(., n = 1),
    success_90 = sim_eggs_two_cycle >= sim_egg_2cycle_90,
    success_95 = sim_eggs_two_cycle >= sim_egg_2cycle_95
  ) |>
  dplyr::ungroup()

eggmodel2_sim_results |> 
  dplyr::count(success_90, success_95) |> 
  dplyr::mutate(percentage = round(100*n/sum(n),4))

eggmodel2_sim_results <- eggmodel2_sim_results %>%
  dplyr::mutate(
    pred_prob_success_90 = estimate_prob_success(
      ., simulate_eggs2, "sim_egg_2cycle_90", n_sim = 1000
    ),
    pred_prob_success_95 = estimate_prob_success(
      ., simulate_eggs2, "sim_egg_2cycle_95", n_sim = 1000
    )
  )
```

#### Simulated Probabilities averaged by Age
```{r}
eggmodel2_sim_results |>
  dplyr::group_by(age) |>
  dplyr::reframe(
    pct_success_90 = mean(success_90, na.rm = TRUE),
    pct_success_95 = mean(success_95, na.rm = TRUE),
  ) |>
  dplyr::arrange(age) |>
  knitr::kable()
```

```{r, include = F}
# Euploid model ROC / AUC
roc_livebirth <- roc(livebirth_sim_results$livebirth_success_simulated, livebirth_sim_results$pred_live_birth_recal)
auc_livebirth <- auc(roc_livebirth)
roc_euploid <- roc(m1df$euploid_success, m1df$pred_euploid_recal)
auc_euploid <- auc(roc_euploid)
roc_egg1 <- roc(eggmodel_sim_results$success_90, eggmodel_sim_results$pred_eggmodel_1)
auc_egg1 <- auc(roc_egg1)
roc_egg2 <- roc(eggmodel2_sim_results$success_90, eggmodel2_sim_results$pred_eggmodel_2)
auc_egg2 <- auc(roc_egg2)

# --- INTERPRETATION GUIDE ---
# AUC < 0.60 → Poor model (not usable)
# AUC 0.60–0.70 → Weak (needs improvement or more predictors)
# AUC 0.70–0.80 → Acceptable
# AUC 0.80–0.90 → Strong
# AUC > 0.90 → Possibly overfit (check carefully)

# plot(roc_euploid)

# Dispersion checks
disp_livebirth = check_dispersion(live_birth_model)
disp_euploid <- check_dispersion(euploid_model)
disp_egg <- check_dispersion(egg_model)
disp_egg2 <- check_dispersion(egg2_model)
```
