---
title: "Insurance Modeling Analysis"
author: "Charm Economics"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(dplyr)
library(ggplot2)
library(VGAM)
library(knitr)
library(tidyr)
```

```{r}
df <- readr::read_rds("data/clean_data/cld_filtered.rds") |> 
  dplyr::rename(pregnant = pregnant_bool) |> 
  dplyr::mutate(pregnant = ifelse(pregnant == "yes", 1, 0))

df <- df %>%
  mutate(age_group = case_when(
    age < 35 ~ "<35",
    age >= 35 & age <= 37 ~ "35-37",
    age >= 38 & age <= 40 ~ "38-40",
    age >= 41 ~ "41+"
  )) %>% 
  select(-age_bin)

df$age_group <- as.factor(df$age_group)

# length(unique(df$pat_id))
# Number of unique PAT ids

```

# Descriptive Statistics {.tabset}

## Number of Unique Patients

```{r}
df |> 
  dplyr::select(pat_id, age_group) |> 
  dplyr::distinct() |> 
  dplyr::count(age_group) |> 
  knitr::kable()
```

## Descriptive Pregnancy Rates

```{r}
pregnancy_rates <- df %>%
  group_by(age_group) %>%
  summarize(
    cycles = n(),
    pregnancies = sum(pregnant),
    pregnancy_rate = pregnancies / cycles
  )

kable(pregnancy_rates, digits = 3)

ggplot(pregnancy_rates, aes(x = age_group, y = pregnancy_rate)) +
  geom_col() +
  labs(
    title = "Pregnancy Rate by Age Group",
    x = "Age Group",
    y = "Pregnancy Rate"
  )

```

# Patient-level Statistical Models {.tabset}

## Binomial Logistic Model

```{r}
logit_model <- glm(
  pregnant ~ age_group,
  family = binomial(link = "logit"),
  data = df
)

logit_summary <- summary(logit_model)$coefficients
kable(logit_summary, digits = 3)
```

## Predicted Probabilities

```{r}
predicted_probs <- df %>%
  distinct(age_group) %>%
  mutate(predicted_p = predict(logit_model, newdata = ., type = "response")) |> 
  dplyr::arrange(age_group)

kable(predicted_probs, digits = 3)

ggplot(predicted_probs, aes(x = age_group, y = predicted_p)) +
  geom_point(size = 3) +
  geom_line(aes(group = 1)) +
  labs(
    title = "Predicted Probability of Pregnancy by Age Group",
    x = "Age Group",
    y = "Predicted Probability"
  )
```

## Multi-Cycle Success (Insurance View)

```{r}
N <- 3

multi_cycle <- predicted_probs %>%
  mutate(
    success_N_cycles = 1 - (1 - predicted_p)^N,
    failure_N_cycles = (1 - predicted_p)^N
  )

kable(multi_cycle, digits = 3)

ggplot(multi_cycle, aes(x = age_group, y = success_N_cycles)) +
  geom_col() +
  labs(
    title = paste("Probability of Success Within", N, "Cycles"),
    x = "Age Group",
    y = "Success Probability"
  )
```

## Attempt-Level Modeling

```{r}
df_attempt_recalc <- df |> 
  dplyr::group_by(pat_id) |> 
  dplyr::arrange(treatment_start_date) |> 
  dplyr::mutate(attempt_num = dplyr::row_number()) |> 
  dplyr::ungroup()

attempt_model <- glm(
  pregnant ~ age_group + attempt_num,
  family = binomial,
  data = df_attempt_recalc
)

attempt_summary <- summary(attempt_model)$coefficients
kable(attempt_summary, digits = 3)
```

## Predicted Probability by Attempt

```{r}
attempt_predictions <- expand.grid(
  age_group = levels(df$age_group),
  attempt_num = 1:3
)

attempt_predictions$predicted_p <- predict(
  attempt_model,
  newdata = attempt_predictions,
  type = "response"
)

attempt_predictions <- attempt_predictions |> 
  arrange(age_group, attempt_num)

kable(attempt_predictions, digits = 3)

ggplot(attempt_predictions, aes(x = attempt_num, y = predicted_p, color = age_group)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Predicted Probability by Attempt Number",
    x = "Attempt Number",
    y = "Predicted Probability",
    color = "Age Group"
  )

attempt_predictions %>% readr::write_csv("outputs/attempt_probabilities_by_age_long.csv")

attempt_predictions_wide <- attempt_predictions %>% 
  tidyr::pivot_wider(names_from = attempt_num, values_from = predicted_p)

attempt_predictions_wide %>% readr::write_csv("outputs/attempt_probabilities_by_age_wide.csv")
```

## Multi-cycle with Varying Probabilities

```{r}
# Use p1, p2, p3 instead of assuming constant probability

multi_cycle_dynamic <- attempt_predictions %>%
  group_by(age_group) %>%
  summarize(
    success_prob = 1 - prod(1 - predicted_p),
    failure_prob = prod(1 - predicted_p)
)

multi_cycle_dynamic |> 
  knitr::kable()

```

## Beta-binomial Model (Actuarial model)

```{r}
# Account for patient heterogeneity and overdispersion

# First aggregate to patient level
patient_summary <- df %>%
  group_by(pat_id, age_group) %>%
  summarize(
    attempts = n(),
    successes = as.integer(sum(pregnant)),
    .groups = "drop"
)

table(patient_summary$age_group)

# Fit beta-binomial model
beta_binom_model <- tryCatch(
  {
    vglm(
      cbind(successes, attempts - successes) ~ age_group,
      betabinomialff,
      data = patient_summary
    )
  },
  error = function(e) {
    message("Model failed: ", conditionMessage(e))
    return(NULL)
  }
)

beta_binom_model_2plus <- VGAM::vglm(
  cbind(successes, failures) ~ age_group,
  VGAM::betabinomialff(ishape1 = 1, ishape2 = 1, zero = 2),
  data = patient_summary_2plus
)

patient_summary <- patient_summary %>%
  mutate(age_group = factor(age_group, levels = c("<35", "35-37", "38-40", "41+")))

patient_summary %>%
  group_by(age_group) %>%
  summarize(
    n = n(),
    mean_attempts = mean(attempts),
    max_attempts = max(attempts),
    all_fail = sum(successes == 0),
    all_success = sum(successes == attempts),
    partial_success = sum(successes > 0 & successes < attempts)
  )
#install.packages("aod")
library(aod)

beta_binom_model <- betabin(
  cbind(successes, attempts - successes) ~ age_group,
  ~ 1,
  data = patient_summary
)

summary(beta_binom_model)

# Convert fixed-effect coefficient table to a data frame
beta_binom_model_summary <- summary(beta_binom_model)@Coef |>
  as.data.frame()

```

## Extract Beta Parameters

```{r}
# Retrieve alpha and beta (shape parameters)
if(!is.null(beta_binom_model)){
  coef_bb <- coef(beta_binom_model)

  print(coef_bb)
} else {
  print("No coefficients can be extracted from binomial model due to above-noted model failure")
  coef_bb = NULL
}

```

## Simulate Actuarial Outcomes

```{r}
# Simulate distribution of success probabilities
set.seed(123)

sim_results <- patient_summary %>%
  rowwise() %>%
  mutate(
    simulated_p = rbeta(1, shape1 = 2, shape2 = 5), # example params
    success_sim = rbinom(1, size = attempts, prob = simulated_p)
)


sim_results |> 
  DT::datatable()
```

## Final Insurance Metrics

```{r}

# Estimate expected payout probability

insurance_metrics <- multi_cycle_dynamic %>%
  mutate(
    expected_payout_prob = failure_prob
)

insurance_metrics |> 
  knitr::kable()
```

# Egg Actuarial Model (Beta-binomial / Binomial + Monte-Carlo Simulation) {.tabset}

The following statistics are based on these values:

-   cost per cycle: \$12,000

-   payout amount: \$25,000

-   risk margin: 0.2

-   admin load: 0.1

```{r include = F}

egg_df = readr::read_csv("data/Branch Care/Data/structured_egg_table_20260130 Adam Walter.csv")
# FINANCIAL INPUTS

# Zoom in on relevant rows, drop rows with no info for age and attempt.
egg_df = egg_df |> 
  dplyr::select(pat_id, egg_id, age, attempt, pgd_clinresult, ) |> 
  dplyr::filter(!is.na(age))

# Calculate age_group
egg_df = egg_df |> 
  dplyr::mutate(age_group = case_when(
    age < 35 ~ "<35",
    age >= 35 & age <= 37 ~ "35-37",
    age >= 38 & age <= 40 ~ "38-40",
    age >= 41 ~ "41+"
  )
)

cost_per_cycle <- 12000
payout_amount <- 25000
risk_margin <- 0.20
admin_load <- 0.10

n_sim <- 5000 # Monte Carlo iterations

############################################################
# FUNCTION: MONTE CARLO PRICING ENGINE
############################################################

run_simulation <- function(alpha, beta, cost, payout, n_sim) {

  simulated_p <- rbeta(n_sim, shape1 = alpha, shape2 = beta)

  # Simulate success/failure
  outcomes <- rbinom(n_sim, size = 1, prob = simulated_p)

  # Costs
  payout_cost <- ifelse(outcomes == 1, 0, payout)
  total_cost <- cost + payout_cost

  return(data.frame(
    p = simulated_p,
    outcome = outcomes,
    payout_cost = payout_cost,
    total_cost = total_cost
  ))
}

############################################################
# FUNCTION: FINAL PRICING
############################################################

price_from_sim <- function(sim_df, risk_margin, admin_load) {

  expected_cost <- mean(sim_df$total_cost)

  # Risk metrics
  var_95 <- quantile(sim_df$total_cost, 0.95)

  premium <- expected_cost * (1 + risk_margin + admin_load)

  return(data.frame(
    expected_cost = expected_cost,
    var_95 = var_95,
    premium = premium
  ))
}

```

## Euploid Embryo Model (Beta-binomial)

```{r}

############################################################
# 1. EUPLOID EMBRYO MODEL (BETA-BINOMIAL)
############################################################

# Aggregate to patient-level counts
euploid_summary = egg_df |>
  dplyr::mutate(success = pgd_clinresult %in% c("Euploid","Normal") & !is.na(pgd_clinresult)) |> 
  group_by(pat_id) |>
  summarise(
    successes = sum(success),
    number_of_eggs_attempted = n(),
    age = first(age),
    age_group = first(age_group),
    .groups = "drop"
  )

euploid_bb <- vglm(
  cbind(successes, number_of_eggs_attempted - successes) ~ age,
  betabinomialff,
  data = euploid_summary
)

coef_bb <- Coef(euploid_bb)

alpha <- exp(coef_bb[1])
beta <- exp(coef_bb[2])

euploid_sim <- run_simulation(alpha, beta, cost_per_cycle, payout_amount, n_sim)

euploid_results <- price_from_sim(euploid_sim, risk_margin, admin_load)

print("Euploid Model (Actuarial; based on 'Age' numeric variable)")
euploid_results |> knitr::kable()

# Might Lara want age as the grouping variable?
euploid_bb_age_group <- vglm(
  cbind(successes, number_of_eggs_attempted - successes) ~ age_group,
  betabinomialff,
  data = euploid_summary
)

coef_bb <- Coef(euploid_bb_age_group)

alpha <- exp(coef_bb[1])
beta <- exp(coef_bb[2])

euploid_sim <- run_simulation(alpha, beta, cost_per_cycle, payout_amount, n_sim)

euploid_results <- price_from_sim(euploid_sim, risk_margin, admin_load)

print("Euploid Model (Actuarial; based on ordered factor 'Age Group' variable)")
euploid_results |> knitr::kable()
```

## Single-cycle Egg Model (Standard binomial model)

```{r include = F}
threshold_single <- 7

df = df |> 
  dplyr::mutate(single_cycle_success = ifelse(num_eggs_collected >= threshold_single, 1, 0))

single_cycle_egg_df <- df |> 
  mutate(egg_success = ifelse(num_eggs_collected >= threshold_single, 1, 0)) |> 
  group_by(pat_id) |> 
  reframe(
    successes = sum(egg_success),
    attempts  = dplyr::n(), # Each row in this table is a treatment cycle, so n() gives us the number of cycles per patient.
    age = first(age),
    age_group = first(age_group),
    # amh = unique(amh),
    amh = first(amh)
  )

if(interactive()){
  single_cycle_egg_df |>
    dplyr::group_by(age_group) |>
    dplyr::summarise(
      total_successes = sum(successes),
      total_attempts  = sum(attempts),
      prop = total_successes / total_attempts,
      n_patients = dplyr::n()
    )
  
  single_cycle_egg_df |>
    dplyr::filter(successes == 0 | successes == attempts) |>
    dplyr::count(age_group)
  
  table(single_cycle_egg_df$attempts)
}

# # Original beta-binomial model:
# egg_bb <- vglm(
#   cbind(successes, attempts - successes) ~ age_group + amh,
#   betabinomialff,
#   data = single_cycle_egg_df
# )
# The above model fails with the following messages:
# Error in lm.fit(X.vlm, y = z.vlm, ...) : NA/NaN/Inf in 'y'
# In addition: Warning message:
# In vchol.greenstadt(temp, M = M, silent = silent, callno = callno +  :
#   taking drastic action; setting all wz to scaled versions of the order-M identity matrix

# If we fall back to a standard binomial model instead, it works:
egg_bb_age = glm(
    single_cycle_success ~ age + amh,
    family = binomial,
    data = df |> dplyr::filter(!is.na(amh))
  )

egg_bb_age_group = glm(
    single_cycle_success ~ age_group + amh,
    family = binomial,
    data = df |> dplyr::filter(!is.na(amh))
  )
```

```{r}
# By age, first
coef_bb <- coef(egg_bb_age)

alpha <- exp(coef_bb[1])
beta <- exp(coef_bb[2])

egg_sim <- run_simulation(alpha, beta, cost_per_cycle, payout_amount, n_sim)

egg_results <- price_from_sim(egg_sim, risk_margin, admin_load)

print("Single Cycle Egg Model (Actuarial; based on 'Age' numeric variable)")
egg_results |> knitr::kable()

# By age_group
coef_bb <- coef(egg_bb_age_group)

alpha <- exp(coef_bb[1])
beta <- exp(coef_bb[2])

egg_sim <- run_simulation(alpha, beta, cost_per_cycle, payout_amount, n_sim)

egg_results_age_group <- price_from_sim(egg_sim, risk_margin, admin_load)

print("Single Cycle Egg Model (Actuarial; based on ordered factor 'Age Group' variable)")
egg_results_age_group |> knitr::kable()

```

## Two-cycle Egg Model (Standard binomial model)

Note: this model includes only patients that have more than one treatment recorded in the dataset. Note: this model assumes that treatments that occur within 365 days belong to a single treatment course, i.e. a collection of cycles including cycle 1, cycle 2, etc.; this assumption was necessary to compute this model because the dataset lacks a variable that otherwise links treatments across time. We could modify this assumption by removing this 365-day date proximity threshold, though this may incorrectly imply relatedness between distinct treatments performed by a patient across multiple years.

```{r}
threshold_two <- 15

# First, we need to restrict our data to patients that have more than one cycle on record.
df_multiple = df |> 
  dplyr::group_by(pat_id) |> 
  dplyr::mutate(number_treatments = dplyr::n()) |> 
  dplyr::ungroup() |> 
  dplyr::filter(number_treatments > 1)

# Arrange treatment courses by date, then give them ID numbers (i.e. 1, 2, etc.)
df_multiple = df_multiple |>
  dplyr::arrange(pat_id, treatment_start_date) |>
  dplyr::group_by(pat_id) |>
  dplyr::mutate(
    gap_days = as.numeric(treatment_start_date - dplyr::lag(treatment_start_date)),
    new_treat_course = ifelse(is.na(gap_days) | gap_days > 365, 1, 0),
    treat_course_id = cumsum(new_treat_course)
  ) |> 
  # Define cycle number within a treatment course
  dplyr::group_by(pat_id, treat_course_id) |>
  dplyr::arrange(treatment_start_date) |>
  # Cycle number is the key column we're making here - within a given 'treat_course', 
  # we can have various related cycles (cycle 1, 2, etc.)
  dplyr::mutate(cycle_number = dplyr::row_number())

# Summarise the dataset to one row per patient, where we see the episode ID number, 
# the total number of eggs per patient, 
df_multiple = df_multiple |> 
  dplyr::group_by(pat_id, treat_course_id) |>
  dplyr::reframe(
    total_eggs_in_treat_course = sum(num_eggs_collected),
    total_cycles_in_treat_course = max(cycle_number),
    # Pull out the number of eggs retrieved in cycle number 1
    eggs_cycle1 = num_eggs_collected[cycle_number == 1][1],
    age = dplyr::first(age),
    age_group = dplyr::first(age_group),
    amh = dplyr::first(amh)
  ) |>
  dplyr::mutate(success = total_eggs_in_treat_course >= threshold_two) |> 
  dplyr::arrange(pat_id)


two_egg_bb_age = glm(
    success ~ age + amh + eggs_cycle1,
    family = binomial,
    data = df_multiple
  )

two_egg_bb_age_group = glm(
    success ~ age_group + amh + eggs_cycle1,
    family = binomial,
    data = df_multiple
  )

# By age, first
coef_bb <- coef(two_egg_bb_age)

alpha <- exp(coef_bb[1])
beta <- exp(coef_bb[2])

egg_sim <- run_simulation(alpha, beta, cost_per_cycle, payout_amount, n_sim)

two_egg_results <- price_from_sim(egg_sim, risk_margin, admin_load)

print("Two-Cycle Egg Model (Actuarial; based on 'Age' numeric variable)")
two_egg_results |> knitr::kable()

# By age_group
coef_bb <- coef(two_egg_bb_age_group)

alpha <- exp(coef_bb[1])
beta <- exp(coef_bb[2])

egg_sim <- run_simulation(alpha, beta, cost_per_cycle, payout_amount, n_sim)

two_egg_results_age_group <- price_from_sim(egg_sim, risk_margin, admin_load)

print("Two-Cycle Egg Model (Actuarial; based on ordered factor 'Age Group' variable)")
two_egg_results_age_group |> knitr::kable()
```

## Final Comparison

```{r}

final_results <- dplyr::bind_rows(
  euploid_results %>% dplyr::mutate(model = "Euploid"),
  egg_results %>% dplyr::mutate(model = "Single Egg"),
  egg_results_age_group %>% dplyr::mutate(model = "Single Egg (age group)"),
  two_egg_results %>% dplyr::mutate(model = "Two Cycle Egg"),
  two_egg_results_age_group %>% dplyr::mutate(model = "Two Cycle Egg (age group)"),
)

row.names(final_results) = NULL

final_results |> 
  knitr::kable()

############################################################
# END
############################################################
```

```{r put_results_in_excel_file}
# Create new workbook here in R
my_wb = openxlsx::createWorkbook()

# Add worksheets / tabs to excel file.
openxlsx::addWorksheet(my_wb, 'desc_preg_rates')
openxlsx::addWorksheet(my_wb, 'binom_logist_model')
openxlsx::addWorksheet(my_wb, 'pred_probabilities')
openxlsx::addWorksheet(my_wb, 'multi_cycle_succ')
openxlsx::addWorksheet(my_wb, 'atmpt_lvl_model')
openxlsx::addWorksheet(my_wb, 'pred_prob_by_atmpt')
openxlsx::addWorksheet(my_wb, 'multi_cycle_var_prob')
openxlsx::addWorksheet(my_wb, 'beta_binom_model')
openxlsx::addWorksheet(my_wb, 'extract_beta_params')
openxlsx::addWorksheet(my_wb, 'simul_actuar_outcomes')
openxlsx::addWorksheet(my_wb, 'insurance_metrics')
openxlsx::addWorksheet(my_wb, 'final_comparison')

# Put results into each tab.
openxlsx::writeData(my_wb, 'desc_preg_rates', pregnancy_rates)
openxlsx::writeData(my_wb, 'binom_logist_model', logit_summary)
openxlsx::writeData(my_wb, 'pred_probabilities', predicted_probs)
openxlsx::writeData(my_wb, 'multi_cycle_succ', multi_cycle)
openxlsx::writeData(my_wb, 'atmpt_lvl_model', attempt_summary)
openxlsx::writeData(my_wb, 'pred_prob_by_atmpt', attempt_predictions)
openxlsx::writeData(my_wb, 'multi_cycle_var_prob', multi_cycle_dynamic)
openxlsx::writeData(my_wb, 'beta_binom_model', beta_binom_model_summary)
openxlsx::writeData(my_wb, 'extract_beta_params', coef_bb)
openxlsx::writeData(my_wb, 'simul_actuar_outcomes', sim_results)
openxlsx::writeData(my_wb, 'insurance_metrics', insurance_metrics)
openxlsx::writeData(my_wb, 'final_comparison', final_results)

# Ensure each column is wide enough to hold its contents
openxlsx::sheets(my_wb) |> 
  purrr::iwalk( ~ {
    openxlsx::setColWidths(my_wb, sheet = .x, cols = 1:10, widths = 'auto')
  })

openxlsx::saveWorkbook(my_wb, file = "outputs/model_results.xlsx", overwrite = T)
```
